---
title: "Guatemala Absorption"
author: "Emily Linebarger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
library(data.table) 
library(ggplot2)
library(knitr)
library(kableExtra)
library(Hmisc)
library(lubridate)
library(wesanderson)
library(gridExtra)
library(scales)

options(scipen=15)

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=8)

save_loc = "J:/Project/Evaluation/GF/resource_tracking/visualizations/deliverables/_GTM 2019 annual report/"
#Source graphing functions
source("C:/Users/elineb/Documents/gf/resource_tracking/analysis/graphing_functions.r")

#Read in absorption data 
gtm = readRDS("C:/Users/elineb/Box Sync/Global Fund Files/GTM/prepped_data/absorption_gtm.rds")
gtm = gtm[grant_period%in%c('2018-2018', '2018-2020', '2019-2021') | (grant=="GTM-T-MSPAS" & grant_period=="2016-2019" & semester=="Semester 5")]

#Paste the grant period on each grant name for viewing purposes 
gtm[, grant:=paste0(grant, " (", grant_period, ")")]

#Assign each grant a distinct color from the "Darjeeling1" palette. 
grants = unique(gtm$grant) 
for (i in 1:length(grants)){
  g = grants[i]
  gtm[grant==g, grant_color:=wes_palette("Darjeeling1")[i]]
}

#Improve the semester label 
gtm[, date_range:=paste0(month(start_date), "/", year(start_date), "-", 
                         month(end_date), "/", year(end_date))]

#Add abbreviated module names. 
all_mods = readRDS("J:/Project/Evaluation/GF/resource_tracking/modular_framework_mapping/all_interventions.rds")
setnames(all_mods, c('module_eng', 'intervention_eng', 'abbrev_mod_eng', 'abbrev_int_eng'), c('gf_module', 'gf_intervention', 'abbrev_mod', 'abbrev_int'))
all_mods = unique(all_mods[, .(gf_module, gf_intervention, disease, abbrev_mod, abbrev_int)])
gtm = merge(gtm, all_mods, by=c('gf_module', 'gf_intervention', 'disease'))

#make sure this merge worked correctly. 
stopifnot(nrow(gtm[is.na(abbrev_int)])==0)

#Break out into smaller datasets. 
by_grant = gtm[, .(budget=sum(budget, na.rm=TRUE), expenditure=sum(expenditure, na.rm=TRUE)), by=c('grant', 'grant_disease', 'start_date', 'date_range', 'grant_color')]
by_grant[, absorption:=(expenditure/budget)*100]
by_grant[, label1:=paste0(round(absorption, 1), "%")]
by_grant[absorption>200, absorption:=200] # Cap absorption at 200%
by_grant[, label2:=paste0(grant_disease)]

by_module = gtm[, .(budget=sum(budget, na.rm=TRUE), expenditure=sum(expenditure, na.rm=TRUE)), by=c('loc_name', 'grant', 'gf_module', 'abbrev_mod', 'start_date', 'date_range', 'grant_color')]
by_module[, absorption:=(expenditure/budget)*100]
by_module[, label1:=paste0(round(absorption, 1), "%")]
by_module[absorption>200, absorption:=200] # Cap absorption at 200%
by_module[, label2:=dollar(expenditure)]

by_intervention = gtm[, .(budget=sum(budget, na.rm=TRUE), expenditure=sum(expenditure, na.rm=TRUE)), by=c('loc_name', 'grant', 'gf_module', 'gf_intervention', 'abbrev_mod', 'abbrev_int', 'start_date', 'date_range', 'grant_color')]
by_intervention[, absorption:=(expenditure/budget)*100]
by_intervention[, label1:=paste0(round(absorption, 1), "%")]
by_intervention[absorption>200, absorption:=200] # Cap absorption at 200%

#Read in budgets file 
budgets = readRDS("C:/Users/elineb/Box Sync/Global Fund Files/GTM/prepped_data/final_budgets.rds")
budgets = merge(budgets, all_mods, by=c('gf_module', 'gf_intervention', 'disease'))
```

## The financial datasets IHME has available are: 
1. Global Fund absorption, created from PUDRs
2. Global Fund final budgets, created from budget files 
3. Global Fund budget revisions, created from budget files (where we have access) 
4. Global Fund expenditure, created from PUDRs
5. Government health expenditure from SICOIN (through June 2018)
6. Other development assistance for health (all other donor spending)

This analysis will use mainly dataset #1 to explore Global Fund absorption.

## Currently available PUDRs 
```{r data_availability} 
by_grant[grant_disease=="malaria", grant_disease:="Malaria"]
by_grant[grant_disease=="hiv", grant_disease:="HIV"]
by_grant[grant_disease=="tb", grant_disease:="TB"]
kable(by_grant[, .(grant, grant_disease, date_range)], col.names=c('Grant', 'Disease', 'PUDR Date Range')) %>%
  kable_styling(font_size=12)
```

I've set up this analysis to include only the most recent PUDR for each grant, because the grant implementation periods vary in Guatemala.   

# 1. Overview of 2018-2019 absorption
```{r overview, fig.height=15} 
p = ggplot(by_grant, aes(x=grant, y=absorption, fill=date_range, label=label1))+ 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(stat="identity", hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  scale_y_continuous(limits=c(0, 220)) + 
  #theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  coord_flip() + 
   scale_fill_manual(values=wes_palette("GrandBudapest2")) + 
  labs(title="Comparison of absorption rates by grant (2018-2019)", x="Grant", y="Absorption (%)", fill="PUDR date range", caption="*Bar height limited to 200%")

ggsave(paste0(save_loc, "absorption_by_grant.png"), p, height=8, width=12)

melt = by_grant[, .(grant, budget, expenditure, label1)]
melt = melt(melt, id.vars=c('grant', 'label1'))
melt[variable=="budget", label1:=""]
melt[variable=="budget", variable:="Budget"]
melt[variable=="expenditure", variable:="Expenditure"]

p2 = ggplot(melt, aes(x=grant, y=value, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label1), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Budget and expenditure by grant"), x="", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")

grid.arrange(p, p2, ncol=1, nrow=2)
```

## Absorption by module for each grant
```{r by_module, results="asis", fig.height=15} 
i = 0
for (g in unique(by_module$grant)){
  subset = by_module[grant==g]
  p = ggplot(subset, aes(x=abbrev_mod, y=absorption, label=label1)) + 
    geom_bar(stat="identity", fill=unique(subset$grant_color)) + 
    theme_bw(base_size=16) + 
    coord_flip() + 
    geom_text(size=4, hjust=0) + 
    scale_y_continuous(limits=c(0, 150)) + 
    labs(title=paste0("Absorption by module for most recent \n", g, " PUDR"), x="Module", y="Absorption (%)")
  # print(p) 
  # if (i<3){cat("\n\n\\pagebreak\n")} 
  
  subset_melt = subset[, .(abbrev_mod, label2, budget, expenditure)]
  subset_melt = melt(subset_melt, id.vars=c('abbrev_mod', 'label2'), variable.name='variable', value.name='amount')
  subset_melt[variable=="budget", label2:=""] #Don't display the expenditure amount on the budget bar. 
  
  #Fix budget and expenditure so they display nicely 
  subset_melt[variable=="budget", variable:="Budget"]
  subset_melt[variable=="expenditure", variable:="Expenditure"]
  
  p2 = ggplot(subset_melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label2), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Budget and expenditure for ", g, "\nby module"), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")
  # print(p2)
  
  ggsave(paste0(save_loc, "budget_exp_", g, ".png"), p2, height=8, width=11)
  grid.arrange(p, p2, ncol=1, nrow=2)
  
  ggsave(paste0(save_loc, g, ".png"), grid.arrange(p, p2, ncol=1, nrow=2), height=15, width=10)
  if (i<3){cat("\n\n\\pagebreak\n")}
  i = i+1
} 
```

## Findings 
Overall, we see a very high proportion of the budget and expenditure for each grant going towards program management. The next step we could take here would be to review what activities they're budgeting for under program management, and talk about what the biggest spending categories are. 

Other than this, one trend we're seeing across countries is that modules with the smallest budget are reporting the highest absorption, and this seems to be the case in Guatemala as well. It's interesting that program management deviates strongly from this trend (grants report both high budget and high absorption under this module). 

Within the HIV grants, we see that human rigths was one of the smaller modules in the HIVOS grant, but it was absorbed at a high level (93.6%). HIVOS also spent most of the money for men who have sex with men and transgender people. On the other hand, the INCAP grant has had trouble spending money for human rights, and absorption for key populations has been mixed. 

We do see high absorption under the "financial management systems" module in the INCAP grant, but there was very little money allocated to this module in the first place. 

Finally, it's important to not that the T-MSPAS and M-MSPAS grants haven't reported any expenditure yet. 

## RSSH absorption across grants 

```{r rssh}
rssh_mods = c('Community systems', 'Financial systems', "Info systems & M&E", "HR & health workers", "Service delivery", "Nat. health strategies")

rssh_by_int = gtm[abbrev_mod%in%rssh_mods, .(budget=sum(budget, na.rm=TRUE), expenditure=sum(expenditure, na.rm=TRUE)), by=c('abbrev_mod', 'abbrev_int')]
rssh_by_int[, absorption:=(expenditure/budget)*100]
rssh_by_int[, label1:=paste0(round(absorption, 1), "%")]
rssh_by_int[absorption>200, absorption:=200] # Cap absorption at 200%

#Shorten a few intervention names. 
rssh_by_int[abbrev_int=="Supportive policy and programmatic environment", abbrev_int:="Supportive policy environment"]

#Sort the graph by module 
rssh_by_int[, module_num:=.GRP, by='abbrev_mod']
rssh_by_int[, int_num:=.GRP, by=c('abbrev_mod', 'abbrev_int')]
rssh_by_int[, rssh_factor:=paste0(module_num, int_num)]

rssh_by_int$int_factor = factor(rssh_by_int$rssh_factor, unique(rssh_by_int$rssh_factor), unique(rssh_by_int$abbrev_int))

p1 = ggplot(rssh_by_int, aes(x=int_factor, y=absorption, fill=abbrev_mod, label=label1))+ 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(hjust=-0.5, size=4) + 
  theme_bw(base_size=14) + 
  scale_fill_manual(values=wes_palette(6, name="Rushmore1", type='continuous')) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title="2018-2019 RSSH absorption by intervention, all grants", subtitle="Interventions grouped by module", x="Intervention", y="Absorption (%)",   fill="Module", caption="*Max bar height set at 200%")
```

```{r rssh_budget_exp, fig.height=15} 
  rssh_melt = rssh_by_int[, .(int_factor, label1, budget, expenditure)]
  rssh_melt = melt(rssh_melt, id.vars=c('int_factor', 'label1'), variable.name='variable', value.name='amount')
  rssh_melt[, label1:=dollar(amount)]
  rssh_melt[variable=="budget", label1:=""] #Don't display the expenditure amount on the budget bar. 
  
  #Fix budget and expenditure so they display nicely 
  rssh_melt[variable=="budget", variable:="Budget"]
  rssh_melt[variable=="expenditure", variable:="Expenditure"]
  
  p2 = ggplot(rssh_melt, aes(x=int_factor, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label1), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar, limits=c(0, 350000)) + 
  labs(title=paste0("RSSH budget and expenditure by intervention"), x="Intervention", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")

  grid.arrange(p1, p2, ncol=1)
  ggsave(paste0(save_loc, "rssh.png"), grid.arrange(p1, p2, ncol=1), height=15, width=10)

```

\newpage 
## Findings 
"Financial management" and "program and data quality" stand out as the modules with the highest absorption, and it's notable that program and data quality was able to achieve this even with a large budget allocation. *Note that "Program and data quality" is part of HMIS. Several important RSSH interventions have not yet recorded any expenditure, namely laboratory systems, surveys, and other HMIS interventions. 

\newpage
# 2. Analyses to support deep dives
## TB laboratory systems and case diagnosis spending 
```{r tb_deep_dive, fig.height=13} 
plot_data = budgets[grant_disease=="tb" & abbrev_int=="Case detection and diagnosis", .(budget=sum(budget, na.rm=TRUE)), by=c('abbrev_mod', 'start_date')]
plot_data[abbrev_mod=="Care and prevention", abbrev_mod:="TB care and prevention"]
p1 = ggplot(plot_data, aes(x=start_date, y=budget, color=abbrev_mod)) + 
  geom_point() + 
  geom_line() + 
  theme_bw() + 
  scale_y_continuous(labels=scales::dollar) + 
  labs(title="Funding for case detection and diagnosis over time", x="Date", y="Module", color="Module")

subset = by_intervention[grant=="GTM-T-MSPAS (2016-2019)" & abbrev_int=="Case detection and diagnosis"]
subset[, label2:=dollar(budget)]
p2 = ggplot(subset, aes(x=abbrev_mod, y=budget, label=label2)) + 
  geom_bar(stat="identity", fill=unique(subset$grant_color)) + 
  theme_bw(base_size=16) + 
  geom_text(size=4, hjust=0) + 
  scale_y_continuous(labels=scales::dollar, limits=c(0, 175000)) + 
  labs(title="Case detection and diagnosis budget for GTM-T-MSPAS \nin most recent PUDR", x="Intervention", y="Budget", caption="*Expenditure was $0 for all modules during this period") + 
  coord_flip() 

grid.arrange(p1, p2, ncol=1)
```


\newpage 
## Malaria spending, both on bednets and otherwise
```{r malaria_budgets} 
mal_budgets = budgets[grant=="GTM-M-MSPAS", .(budget=sum(budget)), by=c('grant', 'abbrev_mod', 'abbrev_int')]
mal_budgets[, label:=dollar(budget)]

p1 = ggplot(mal_budgets, aes(x=abbrev_int, y=budget, label=label)) + 
  geom_bar(stat="identity", fill="forestgreen") + 
  theme_bw(base_size=16) + 
  geom_text(size=4, hjust=0) + 
  scale_y_continuous(labels=scales::dollar, limits=c(0, 275000)) + 
  coord_flip() + 
  labs(title="Budget for malaria interventions from 2014-2021", x="Intervention", y="Budget", caption="*Totals show budget for entire grant period") 
```

```{r bednets_over_time, fig.height=14} 
bednet_ints = c("LLIN: Continuous distribution", "LLIN: Mass campaign") 
bednet_budgets = budgets[abbrev_int%in%bednet_ints, .(budget=sum(budget, na.rm=TRUE)), by='start_date']

p2 = ggplot(bednet_budgets, aes(x=start_date, y=budget)) + 
  geom_point(color="hotpink4") + 
  geom_line(color="hotpink4") + 
  theme_bw(base_size=16) + 
  scale_y_continuous(label=scales::dollar) +
  labs(title="Global Fund investment in bednets over time", x="Date", y="Budget")

grid.arrange(p1, p2, ncol=1)
```

## HIV Testing Absorption - to support SIGSA/SIGPRO analyses 
```{r sigsa1} 
#Make funding landscape graph 
funding_landscape("ribbon", "gtm", "hiv", 2010, 2017, altCaption = "*Other DAH data only available until 2017", includeGHE=TRUE)
```

```{r sigsa2}
# What did INCAP spend by module? 
p2 = absorption_by_loc_disease("gtm", "hiv", "2018-2020", byModule=TRUE, grantName="GTM-H-INCAP", barLabels=TRUE)
```

```{r sigsa3, fig.height=13}
#Within key populations, what % went to testing? 
dt = readRDS("C:/Users/elineb/Box Sync/Global Fund Files/GTM/prepped_data/absorption_gtm.rds")
testing_codes = c("H2_7", "H3_7", "H4_6", "H5_7", "H6_7", "H7_3", "H8_5", "H14", "H14_1")
test_data = dt[code%in%testing_codes & grant%in%c('GTM-H-HIVOS', 'GTM-H-INCAP')]

#Fix a few intervention names 
test_data[gf_intervention=="HIV testing services for transgender people", gf_intervention:="Transgender people"]
test_data[gf_intervention=="HIV testing services for men who have sex with men", gf_intervention:="Men who have sex with men"]
test_data[gf_intervention=="Differentiated HIV testing services", gf_intervention:="Differentiated testing"]

p3 = ggplot(test_data, aes(x=gf_intervention, y=absorption, fill=grant, label=paste0(round(absorption, 1), "%"))) + 
  geom_bar(stat="identity") + 
  geom_text(size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_fill_manual(values=ihme_divergent[2:3]) + 
  facet_wrap(~grant) + 
  scale_y_continuous(limits=c(0, 100)) + 
  labs(title="2018 absorption for testing interventions by grant", x="Intervention", y="Absorption (%)", fill="Grant")

grid.arrange(p2, p3, ncol=1)

```

One finding here is that even though there is funding for several key populations in the INCAP grant, not all of them are being targeted for HIV testing. INCAP only budgeted for MSM- and transgender people-specific testing, and they've only spent money on testing for men who have sex with men so far. 

## Funding landscape graphs - can support STC conclusions in Enrique's financial analysis
```{r funding_landscape1} 
funding_landscape("proportion", "gtm", "hiv", 2010, 2017, altCaption = "*Other DAH data only available until 2017", includeGHE=TRUE)
```

```{r funding_landscape2} 
p1 = funding_landscape("proportion", "gtm", "tb", 2010, 2017, altCaption = "*Other DAH data only available until 2017", includeGHE=TRUE)
```

```{r funding_landscape3, fig.height=14} 
p2 = funding_landscape("proportion", "gtm", "malaria", 2010, 2017, altCaption = "*Other DAH data only available until 2017", includeGHE=TRUE)

grid.arrange(p1, p2, ncol=1)
```

For all three diseases, it's clear that government health expenditure makes up the majority of funding, and that this amount has been increasing over time. This conclusion is especially pronounced for TB and malaria, where the Global Fund and the government are the only major donors. 
