---
title: "Guatemala budget revisions"
author: "Emily Linebarger"
date: "October 10, 2019"
output: pdf_document
classoption: landscape
---

```{r setup, echo=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(readxl)
library(scales)
options(scipen=100)

dt = readRDS("C:/Users/elineb/Box Sync/Global Fund Files/GTM/prepped_data/budget_revisions.rds") #This is at the intervention, year, quarter level. 
merge1 = dt[, sum(v0, na.rm=T)]
rows1 = nrow(dt)
setnames(dt, c('gf_module', 'gf_intervention'), c('module_eng', 'intervention_eng'))
all_mods = readRDS("J:/Project/Evaluation/GF/resource_tracking/modular_framework_mapping/all_interventions.rds")
all_mods = unique(all_mods[, .(module_eng, intervention_eng, abbrev_mod_eng, abbrev_int_eng)])

# Why are we getting row expansion here? - EMILY ADD THIS VERIFICATION TO PREP MODULE MAP CODE. 
all_mods[, concat1:=paste0(module_eng, intervention_eng)]
all_mods[, concat2:=paste0(abbrev_mod_eng, abbrev_int_eng)]
all_mods[, grp1:=seq(0, 2, by=1), by='concat1']
dt = merge(dt, all_mods, by=c('module_eng', 'intervention_eng'))
merge2 = dt[, sum(v0, na.rm=T)]
rows2 = nrow(dt)
stopifnot(merge1==merge2 & rows1==rows2)

#Collapse to module level. 
by_module = dt[, .(v0=sum(v0, na.rm=T), v1=sum(v1, na.rm=T), v2=sum(v2, na.rm=T), v3=sum(v3, na.rm=T)), by=c('grant', 'grant_period', 'abbrev_mod_eng')]
by_module = melt(by_module, id.vars=c('grant', 'grant_period', 'abbrev_mod_eng'), value.name="budget", variable.name="version")
by_module[, label:=paste0("$", round(budget))]
by_module[, version:=factor(version, levels=c('v3', 'v2', 'v1', 'v0'), labels=c('v3', 'v2', 'v1', 'v0'))] #Reverse this factoring so it shows up on the graph in the right order. 

#Collapse by intervention. 
by_intervention = dt[, .(v0=sum(v0, na.rm=T), v1=sum(v1, na.rm=T), v2=sum(v2, na.rm=T), v3=sum(v3, na.rm=T)), by=c('grant', 'grant_period', 'abbrev_mod_eng', 'abbrev_int_eng')]
by_intervention = melt(by_intervention, id.vars=c('grant', 'grant_period', 'abbrev_mod_eng', 'abbrev_int_eng'), value.name="budget", variable.name="version")
by_intervention[, label:=paste0("$", round(budget))]
by_intervention[, version:=factor(version, levels=c('v3', 'v2', 'v1', 'v0'), labels=c('v3', 'v2', 'v1', 'v0'))] #Reverse this factoring so it shows up on the graph in the right order. 

#Read in the master file list to pull the names of the file revisions. 
file_list = data.table(read_excel("C:/Users/elineb/Box Sync/Global Fund Files/master_file_list.xlsx"))

```

## List of budget revisions for Guatemala
```{r list_files, echo=FALSE, warning=FALSE}
revision_concats = unique(dt[, paste0(grant, "_", grant_period)])
file_list = unique(file_list[paste0(grant, "_", grant_period)%in%revision_concats & data_source=="budget" & 
                      file_iteration%in%c('final', 'revision'), .(grant, grant_period, file_name, update_date)])
file_list[, update_date:=as.Date(update_date, format="%Y-%m-%d")]
file_list = file_list[order(grant, grant_period, update_date)]

# Add an 'order' variable
file_list[, revision_order:=seq(0, 5, by=1), by=c('grant', 'grant_period')]
file_list[, version:=paste0("version ", revision_order)]
file_list$revision_order <- NULL
kable(file_list[, 1:3])
kable(file_list[, 4:5])
```

## GTM-M-MSPAS
```{r module_graphs1, echo=FALSE, warning=FALSE}
for (g in unique(by_module$grant)){
  p = ggplot(by_module[grant==g], aes(x=abbrev_mod_eng, y=budget, fill=version)) + 
    geom_bar(stat="identity", position="dodge") + 
    # geom_text(data=by_module, aes(x=abbrev_mod_eng, y=budget, label=label), vjust=0) + 
    theme_bw() + 
    coord_flip() + 
    labs(title=paste0(g, " budget revisions, by module"), x="Module", y="Budget", fill="Version") + 
    scale_y_continuous(labels = scales::dollar)
  print(p)
}
```

So far, the only grant we've received updated budgets for is GTM-M-MSPAS. Overall, it looks like the shifts within modules have been minimal, but the largest ones have been within procurement and supply chain management, info systems & M&E, and case management. 

## Procurement and supply chain management budget revisions
```{r by_int1, echo=FALSE, warning=FALSE}
by_intervention[abbrev_int_eng=="Policy, planning, and coordination", abbrev_int_eng:="Policy, planning, \nand coordination"]
by_intervention[abbrev_int_eng=="National disease control programs", abbrev_int_eng:="National disease \ncontrol programs"]

ggplot(by_intervention[abbrev_mod_eng=="PSM"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("PSM budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

This is the intervention-level breakdown of the PSM module. About $150,000 has been taken away from "Supply chain infrastructure and development of tools". 

## Info systems and M&E budget revisions
```{r by_int2, echo=FALSE, warning=FALSE}
ggplot(by_intervention[abbrev_mod_eng=="Info systems & M&E"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("Info systems budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

This slide is a breakdown of the interventions under the module "Health management information systems and M&E." There has been some money added to all interventions under the second version. 

## Case management budget revisions
```{r by_int3, echo=FALSE, warning=FALSE}
ggplot(by_intervention[abbrev_mod_eng=="Case management"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("Case management budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

This slide is a breakdown of the interventions under the module "Case management."All case management interventions have decreased slightly, except for IECC/BCC case management, which has increased. 