---
title: "Uganda 2018-2020 budget revisions"
author: "Emily Linebarger"
date: "October 11, 2019"
output: pdf_document
classoption: landscape
---

```{r setup, echo=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(readxl)
library(scales)
options(scipen=100)

dt = readRDS("C:/Users/elineb/Box Sync/Global Fund Files/UGA/prepped_data/budget_revisions.rds") #This is at the intervention, year, quarter level. 
merge1 = dt[, sum(v0, na.rm=T)]
rows1 = nrow(dt)
setnames(dt, c('gf_module', 'gf_intervention'), c('module_eng', 'intervention_eng'))
all_mods = readRDS("J:/Project/Evaluation/GF/resource_tracking/modular_framework_mapping/all_interventions.rds")
all_mods = unique(all_mods[, .(module_eng, intervention_eng, abbrev_mod_eng, abbrev_int_eng)])

# Why are we getting row expansion here? - EMILY ADD THIS VERIFICATION TO PREP MODULE MAP CODE. 
all_mods[, concat1:=paste0(module_eng, intervention_eng)]
all_mods[, concat2:=paste0(abbrev_mod_eng, abbrev_int_eng)]
all_mods[, grp1:=seq(0, 2, by=1), by='concat1']
dt = merge(dt, all_mods, by=c('module_eng', 'intervention_eng'))
merge2 = dt[, sum(v0, na.rm=T)]
rows2 = nrow(dt)
stopifnot(merge1==merge2 & rows1==rows2)

#Collapse to module level. 
by_module = dt[, .(v0=sum(v0, na.rm=T), v1=sum(v1, na.rm=T), v2=sum(v2, na.rm=T)), by=c('grant', 'grant_period', 'abbrev_mod_eng')]
by_module = melt(by_module, id.vars=c('grant', 'grant_period', 'abbrev_mod_eng'), value.name="budget", variable.name="version")
by_module[, label:=paste0("$", round(budget))]
by_module[, version:=factor(version, levels=c('v2', 'v1', 'v0'), labels=c('v2', 'v1', 'v0'))] #Reverse this factoring so it shows up on the graph in the right order. 

#Collapse to intervention level. 
by_intervention = dt[, .(v0=sum(v0, na.rm=T), v1=sum(v1, na.rm=T), v2=sum(v2, na.rm=T)), by=c('grant', 'grant_period', 'abbrev_mod_eng', 'abbrev_int_eng')]
by_intervention = melt(by_intervention, id.vars=c('grant', 'grant_period', 'abbrev_mod_eng', 'abbrev_int_eng'), value.name="budget", variable.name="version")
by_intervention[, label:=paste0("$", round(budget))]
by_intervention[, version:=factor(version, levels=c('v2', 'v1', 'v0'), labels=c('v2', 'v1', 'v0'))] #Reverse this factoring so it shows up on the graph in the right order. 

#Read in the master file list to pull the names of the file revisions. 
file_list = data.table(read_excel("C:/Users/elineb/Box Sync/Global Fund Files/master_file_list.xlsx"))

```

# List of budget revisions for Uganda
```{r list_files, echo=FALSE, warning=FALSE}
revision_concats = unique(dt[, paste0(grant, "_", grant_period)])
file_list = unique(file_list[paste0(grant, "_", grant_period)%in%revision_concats & data_source=="budget" & 
                      file_iteration%in%c('final', 'revision'), .(grant, grant_period, file_name, update_date)])
file_list[, update_date:=as.Date(update_date, format="%Y-%m-%d")]
file_list = file_list[order(grant, grant_period, update_date)]

# Add an 'order' variable
file_list[, revision_order:=seq(0, 5, by=1), by=c('grant', 'grant_period')]
file_list[, version:=paste0("version ", revision_order)]
file_list$revision_order <- NULL
kable(file_list[, 1:3])
kable(file_list[, 4:5])
```

# UGA-C-TASO module-level budget revisions
```{r C_TASO1, echo=FALSE, warning=FALSE}
ggplot(by_module[grant=="UGA-C-TASO" & version!="v2"], aes(x=abbrev_mod_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  # geom_text(data=by_module, aes(x=abbrev_mod_eng, y=budget, label=label), vjust=0) + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("UGA-C-TASO budget revisions, by module"), x="Module", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

The biggest changes here are to program management, prevention programs for youth/adolescents, MDR-TB, and human rights barriers, so let's review each module by intervention. 

## Revisions to "Program management" for UGA-C-TASO
```{r C_TASO2, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="UGA-C-TASO" & version!="v2" & abbrev_mod_eng=="Program mgmt"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("UGA-C-TASO program management \nbudget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

More money has been allocated to grant management. 


## Revisions to "Youth/adolescent prevention" for UGA-C-TASO
```{r C_TASO3, echo=FALSE, warning=FALSE}
by_intervention[abbrev_int_eng=="Gender-based violence programming for adolescents and youth", abbrev_int_eng:="Gender-based violence"]
by_intervention[abbrev_int_eng=="Behavioral interventions for adolescents and youth", abbrev_int_eng:="Behavioral interventions"]
by_intervention[abbrev_int_eng=="Addressing stigma against youth and adolescents", abbrev_int_eng:="Addressing stigma"]
by_intervention[abbrev_int_eng=="Other interventions adolescents and youth", abbrev_int_eng:="Other interventions"]
ggplot(by_intervention[grant=="UGA-C-TASO" & version!="v2" & abbrev_mod_eng=="Prevention programs for youth/adol."], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("UGA-C-TASO youth/adolescent prevention \nbudget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

## Revisions to "Human rights barriers" for UGA-C-TASO
```{r C_TASO4, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="UGA-C-TASO" & version!="v2" & abbrev_mod_eng=="Human rights barriers"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("UGA-C-TASO human rights \nbudget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

# UGA-H-MoFPED 
```{r H_MOFPED1, echo=FALSE, warning=FALSE}
ggplot(by_module[grant=="UGA-H-MoFPED"], aes(x=abbrev_mod_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("UGA-H-MoFPED budget revisions, by module"), x="Module", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

The largest changes have been to the "Prevention programs for the general population" and "HIV testing services" modules, which are broken down to the intervention-level on the next two pages.  

## Revisions to "Prevention programs for general population" for UGA-H-MoFPED
```{r H_MOFPED2, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="UGA-H-MoFPED" & abbrev_mod_eng=="Prevention programs for general pop."], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("General population prevention \nbudget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

## Revisions to "HIV testing" for UGA-H-MoFPED
```{r H_MOFPED3, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="UGA-H-MoFPED" & abbrev_mod_eng=="HIV Testing Services"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("HIV testing budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

# UGA-T-MoFPED
```{r T_MOFPED1, echo=FALSE, warning=FALSE}
ggplot(by_module[grant=="UGA-T-MoFPED" & version!="v2"], aes(x=abbrev_mod_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("UGA-T-MoFPED budget revisions, by module"), x="Module", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

For T-MoFPED, the changes have been in the MDR-TB and Care & prevention modules. 

## Revisions to "MDR-TB" module for UGA-T-MoFPED
```{r T_MOFPED2, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="UGA-T-MoFPED" & abbrev_mod_eng=="MDR-TB" & version!="v2"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("UGA-T-MoFPED MDR-TB budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

## Revisions to "Care and prevention" module for UGA-T-MoFPED
```{r T_MOFPED3, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="UGA-T-MoFPED" & abbrev_mod_eng=="Care & prevention" & version!="v2"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("UGA-T-MoFPED care & prevention budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```