---
title: "Research Team 3.22.19"
author: "Emily Linebarger"
date: "March 22, 2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
user = "elineb" #Change to your username 
code_dir = ifelse(Sys.info()[1]=='Windows', paste0("C:/Users/", user, "/Documents/gf/"), paste0('/homes/', user, '/gf/'))
source(paste0(code_dir, "resource_tracking/prep/_common/set_up_r.R"), encoding="UTF-8")
library(ggplot2)
dt = readRDS(paste0(fgh_prepped, "ghe_actuals_malaria.rds"))
dt[, value:=as.numeric(value)]
drc = dt[ihme_loc_id == 'COD']
drc[value_code == 'the_malaria', group:='the']
drc[value_code%in%c('fs_malaria_domestic_public', 'fs_malaria_domestic_private_oop', 'fs_malaria_domestic_private'), group:='fs_malaria']
drc[value_code%in%c('func_malaria_cure_care', 'func_malaria_pharm', 'func_malaria_prev', 'func_malaria_complement'), group:='func_malaria']

drc_ghe = drc[value_code == 'fs_malaria_domestic_public']
```

## Exploratory Analysis of GHE spending on malaria from FGH team 

Plot of government health expenditure on malaria 

```{r, echo=FALSE }

# ggplot(drc, aes(x=year_id, y=value, color=source_type, shape=value_code)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Government Health Expenditure on Malaria, DRC") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))
ggplot(drc[value_code=='fs_malaria_domestic_public'], aes(x=year_id, y=value, color=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (fs_malaria_domestic_public)") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))
# ggplot(drc[value_code=='func_malaria_cure_care'], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (func_malaria_cure_care)")
# ggplot(drc[value_code=='func_malaria_fprev'], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (func_malaria_prev)")

ggplot(drc[value_code=='fs_malaria_domestic_public'], aes(x=year_id, y=value, color=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (fs_malaria_domestic_public)") + geom_smooth(method="lm") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))

# ggplot(drc_ghe, aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (fs_malaria_domestic_public)") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))
```
    
    
What unique values do we have in the data, and what years do we have NHA for? 
```{r, echo=FALSE}
drc_ghe[, .(source_type, value, year_id)][order(year_id, source_type)]
drc_ghe[source_type=='nha', .(source_type, value, year_id)]
```

## Questions
Which data source do we want to rely on? Do we want to smooth this trendline?   

How are we handling the conflicting values in 2013 and 2014, where there are outliers in the NHA?    

Should we compare what the FGH team has done with our own PNCNS reports? All of the reports we have are not entered in the FGH team's data.    

Are we comfortable assuming that spending in 2011 was practically 0, or do we need to seek more data?   

Are we okay patching different data sources to make a complete time series? Which ones? 

## Data sources for OOP spending

```{r, echo=FALSE}
drc_oop = drc[value_code=='fs_malaria_domestic_private_oop']
ggplot(drc_oop, aes(x=year_id, y=value, color=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (fs_malaria_domestic_public)") 
drc_oop[, .(year_id, source_type, value)][order(year_id)]
```

```{r, echo=FALSE, include=FALSE}
ggplot(drc, aes(x=year_id, y=value, color=group, shape=group)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Comparison of different data sources (FS, FUNC, and THE)") + geom_line()
ggplot(drc, aes(x=year_id, y=value, color=group, shape=group)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Comparison of different data sources (FS, FUNC, and THE) without trend line") 
ggplot(drc, aes(x=year_id, y=value, color=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Government Health Expenditure on Malaria, DRC") + facet_wrap(vars(group))
#It seems there are three groups of these categories to plot separately: the_malaria (on its own), the three fs_malaria variables and then the func_malaria variables. Could you just make a quick time series plot like the first one, facetted by those three groups (with scales=’free_y’). Might also help to connect the dots.


```


```{r, echo=FALSE, include=FALSE}
unique(drc[, .(year_id, value_code)])[order(year_id)]
# unique(drc[, .(year_id, source_type)])[order(year_id)]
```


```{r, echo=FALSE, include=FALSE}

ggplot(drc_ghe[source_type == "nha"], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, NHAs only") + geom_smooth(method="lm") + 
  scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))
ggplot(drc_ghe[source_type == "gfatm p" | source_type == "gfatm c"], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, Global Fund sources") + geom_smooth(method="lm") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))

ggplot(drc_ghe[source_type == "WMR"], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, other data sources") + geom_smooth(method="lm") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))



```