---
title: "Uganda 2019 Absorption"
author: "Emily Linebarger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
classoption: landscape
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, fig.dim=c(10, 7))

library(data.table) 
library(ggplot2)
library(knitr)
library(kableExtra)
library(Hmisc)
library(wesanderson)
library(gridExtra)
library(scales)
library(RColorBrewer)

dt = readRDS("C:/Users/elineb/Box Sync/Global Fund Files/UGA/prepped_data/absorption_uga.rds")
dt = dt[grant_period=="2018-2020"]

#Assign each grant a distinct color from the "Zissou" palette. 
grants = unique(dt$grant) 
for (i in 1:length(grants)){
  g = grants[i]
  dt[grant==g, grant_color:=wes_palette("Zissou1")[i]]
}

#Add abbreviated module names. 
all_mods = readRDS("J:/Project/Evaluation/GF/resource_tracking/modular_framework_mapping/all_interventions.rds")
setnames(all_mods, c('module_eng', 'intervention_eng', 'abbrev_mod_eng', 'abbrev_int_eng'), c('gf_module', 'gf_intervention', 'abbrev_mod', 'abbrev_int'))
all_mods = all_mods[, .(gf_module, gf_intervention, disease, abbrev_mod, abbrev_int)]
dt = merge(dt, all_mods, by=c('gf_module', 'gf_intervention', 'disease'), allow.cartesian=TRUE)

#make sure this merge worked correctly. 
stopifnot(nrow(dt[is.na(abbrev_int)])==0)

#Break out into smaller datasets. 
by_grant = dt[, .(budget=sum(budget, na.rm=T), expenditure=sum(expenditure, na.rm=T)), by=c('grant', 'grant_disease', 'start_date', 'semester', 'grant_color')]
by_grant[, absorption:=(expenditure/budget)*100]
by_grant[, label1:=paste0(round(absorption, 1), "%")]
by_grant[absorption>200, absorption:=200] # Cap absorption at 200%
by_grant[, label2:=paste0(grant_disease)]

by_module = dt[, .(budget=sum(budget, na.rm=T), expenditure=sum(expenditure, na.rm=T)), by=c('loc_name', 'grant', 'gf_module', 'abbrev_mod', 'start_date', 'semester', 'grant_color')]
by_module[, absorption:=(expenditure/budget)*100]
by_module[, label1:=paste0(round(absorption, 1), "%")]
by_module[absorption>200, absorption:=200] # Cap absorption at 200%
by_module[, label2:=dollar(expenditure)]

by_intervention = dt[, .(budget=sum(budget, na.rm=T), expenditure=sum(expenditure, na.rm=T)), by=c('loc_name', 'grant', 'gf_module', 'gf_intervention', 'abbrev_mod', 'abbrev_int', 'start_date', 'semester', 'grant_color')]
by_intervention[, absorption:=(expenditure/budget)*100]
by_intervention[, label1:=paste0(round(absorption, 1), "%")]
by_intervention[absorption>200, absorption:=200] # Cap absorption at 200%
```

\newpage
# Overview of 2019 absorption
\newpage
```{r overview} 
ggplot(by_grant, aes(x=grant, y=absorption, fill=semester, label=label1))+ 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(stat="identity", position = position_dodge(width = 1), vjust=-0.5, size=4) + 
  theme_bw(base_size=16) + 
  scale_y_continuous(limits=c(0, 220)) + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
   scale_fill_manual(values=wes_palette("GrandBudapest2")) + 
  labs(title="Comparison of 2018-2020 absorption rates by semester", x="Grant", y="Absorption (%)", fill="Grant Semester", caption="*Semester 1 is Q1-Q2 2018, Semester 1-2 is Q1-Q4 2018, and Semester 3 is Q1-Q2 2019\n*Bar height limited to 200%")

```

```{r cumulative}
cumulative = by_grant[, .(grant, grant_disease, semester, budget, expenditure)]
cumulative = melt(cumulative, id.vars=c('grant', 'grant_disease', 'semester'))
cumulative = dcast(cumulative, grant+grant_disease+variable~semester)

#Calculate cumulative budget and expenditure, and then absorption percentage. 
cumulative[!is.na(`Semester 2`), cumulative:=sum(`Semester 1`, `Semester 2`, `Semester 3`), by=c('grant', 'grant_disease', 'variable')]
cumulative[!is.na(`Semester 1-2`), cumulative:=sum(`Semester 1-2`, `Semester 3`), by=c('grant', 'grant_disease', 'variable')]
stopifnot(nrow(cumulative[is.na(cumulative)])==0)

#Subset this so it's easier to work with, and then graph. 
cumulative_by_grant = cumulative[, .(grant, grant_disease, variable, cumulative)]
cumulative_by_grant = dcast(cumulative_by_grant, grant+grant_disease~variable)
cumulative_by_grant[, absorption:=(expenditure/budget)*100]
cumulative_by_grant[, label1:=paste0(round(absorption, 1), "%")]
cumulative_by_grant[absorption>200, absorption:=200] # Cap absorption at 200%
cumulative_by_grant[, label2:=paste0(grant_disease)]

ggplot(cumulative_by_grant, aes(x=grant, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="indianred3") + 
  geom_text(stat="identity", vjust=-0.5, size=4) + 
  theme_bw(base_size=16) + 
  theme(legend.position = "none") +
  scale_y_continuous(limits=c(0, 100)) + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
   scale_fill_manual(values=wes_palette("GrandBudapest2")) + 
  labs(title="Cumulative 2018-2020 absorption rates", x="Grant", y="Absorption (%)", subtitle="(Semesters 1-3)")

#Show cumulative budget and expenditure bars
  cumulative_melt = cumulative_by_grant[, .(grant, budget, expenditure)]
  cumulative_melt = melt(cumulative_melt, id.vars=c('grant'), variable.name='variable', value.name='amount')
  cumulative_melt[variable=="expenditure", label1:=dollar(amount)] #Don't display the expenditure amount on the budget bar. 
  cumulative_melt[is.na(label1), label1:=""]
  
  #Fix budget and expenditure so they display nicely 
  cumulative_melt[variable=="budget", variable:="Budget"]
  cumulative_melt[variable=="expenditure", variable:="Expenditure"]
  
ggplot(cumulative_melt, aes(x=grant, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label1), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_fill_manual(values=c('plum2', 'chocolate2')) + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Cumulative budget and expenditure, by grant"), x="Grant", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")


```

\newpage 
# Supporting data for grant-level absorption
```{r supporting_data1} 
kable(by_grant[, .(grant, grant_disease, start_date, semester, budget, expenditure, absorption)][order(grant, grant_disease, start_date)], col.names=c('Grant', 'Disease', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="c")
```

\newpage 

# 2019 absorption by module/grant 

\newpage

```{r by_module, results='asis'}
i = 0
for(g in unique(by_module$grant)){

  subset = by_module[grant==g & start_date=="2019-01-01"]
  
  p1 = ggplot(subset, aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill=unique(subset$grant_color)) + 
  geom_text(hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Absorption for ", g, " by module"), subtitle="Labels show uncapped absorption percentage", x="Module", y="Absorption (%)", caption="*Max bar height set at 200%")
  print(p1) 
  if (i<4){cat("\n\n\\pagebreak\n")} 
  
  subset_melt = subset[, .(abbrev_mod, label2, budget, expenditure)]
  subset_melt = melt(subset_melt, id.vars=c('abbrev_mod', 'label2'), variable.name='variable', value.name='amount')
  subset_melt[variable=="budget", label2:=""] #Don't display the expenditure amount on the budget bar. 
  
  #Fix budget and expenditure so they display nicely 
  subset_melt[variable=="budget", variable:="Budget"]
  subset_melt[variable=="expenditure", variable:="Expenditure"]
  
  p2 = ggplot(subset_melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label2), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Budget and expenditure for ", g, "\nby module"), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")
  print(p2) 
  if (i<4){cat("\n\n\\pagebreak\n")} 
  
 cumulative = by_module[grant==g, .(grant, semester, budget, expenditure, abbrev_mod)]
cumulative = melt(cumulative, id.vars=c('grant', 'semester', 'abbrev_mod'))
cumulative = dcast(cumulative, grant+abbrev_mod+variable~semester)

#Calculate cumulative budget and expenditure, and then absorption percentage. 
if ("Semester 2"%in%names(cumulative)){
  cumulative[, cumulative:=sum(`Semester 1`, `Semester 2`, `Semester 3`, na.rm=T), by=c('grant', 'abbrev_mod', 'variable')]
} else { 
  cumulative[, cumulative:=sum(`Semester 1-2`, `Semester 3`, na.rm=T), by=c('grant', 'abbrev_mod', 'variable')]
}
stopifnot(nrow(cumulative[is.na(cumulative)])==0)

#Subset this so it's easier to work with, and then graph. 
cumulative_by_module = cumulative[, .(grant, abbrev_mod, variable, cumulative)]
cumulative_by_module = dcast(cumulative_by_module, grant+abbrev_mod~variable)
cumulative_by_module[, absorption:=(expenditure/budget)*100]
cumulative_by_module[, label1:=paste0(round(absorption, 1), "%")]
cumulative_by_module[absorption>200, absorption:=200] # Cap absorption at 200%
cumulative_by_module[, label2:=paste0(abbrev_mod)]

p3 = ggplot(cumulative_by_module, aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="indianred3") + 
  geom_text(hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Cumulative absorption for ", g, "\nby module (Semesters 1-3)"), subtitle="Labels show uncapped absorption percentage", x="Module", y="Absorption (%)", caption="*Max bar height set at 200%")
print(p3)
 if (i<4){cat("\n\n\\pagebreak\n")} 

melt = cumulative_by_module[, .(abbrev_mod, budget, expenditure)]
  melt = melt(melt, id.vars=c('abbrev_mod'), variable.name='variable', value.name='amount')
  melt[variable=="budget", label:=""] #Don't display the expenditure amount on the budget bar. 
  melt[variable=="expenditure", label:=dollar(amount)]
  
  #Fix budget and expenditure so they display nicely 
  melt[variable=="budget", variable:="Budget"]
  melt[variable=="expenditure", variable:="Expenditure"]
  
 p4 =  ggplot(melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_fill_manual(values=c('plum2', 'chocolate2')) + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Cumulative budget and expenditure for \n", g), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")
 print(p4)
 if (i<4){cat("\n\n\\pagebreak\n")} 


  i = i+1
}

```

\newpage
# Supporting data for graphs by grant and module (1/3) 
```{r supporting_data_mod1} 
kable(by_module[1:36, .(grant, abbrev_mod, start_date, semester, budget, expenditure, absorption)][order(grant, start_date, semester, abbrev_mod)], col.names=c('Grant', 'Module', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```

\newpage
# Supporting data for graphs by grant and module (2/3) 
```{r supporting_data_mod2} 
kable(by_module[37:73, .(grant, abbrev_mod, start_date, semester, budget, expenditure, absorption)][order(grant, start_date, semester, abbrev_mod)], col.names=c('Grant', 'Module', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```

\newpage
# Supporting data for graphs by grant and module (3/3) 
```{r supporting_data_mod3} 
stopifnot(nrow(by_module)==108)
kable(by_module[74:108, .(grant, abbrev_mod, start_date, semester, budget, expenditure, absorption)][order(grant, start_date, semester, abbrev_mod)], col.names=c('Grant', 'Module', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```

\newpage
# Absorption by module, pooled by disease
\newpage 

```{r by_module_combined1, results="asis"}
subset_hiv = dt[grant_disease%in%c('hiv', 'hiv/tb'), .(expenditure=sum(expenditure, na.rm=T), budget=sum(budget, na.rm=T)), by=c('semester', 'start_date', 'abbrev_mod')]
subset_hiv[, absorption:=(expenditure/budget)*100]
subset_hiv[, label1:=paste0(round(absorption, 1), "%")]
subset_hiv[absorption>200, absorption:=200] # Cap absorption at 200%
subset_hiv[, label2:=dollar(expenditure)]

ggplot(subset_hiv[start_date=="2019-01-01"], aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="chartreuse4") + 
  geom_text(size=4, hjust=0) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Absorption by module, for HIV and HIV/TB grants"), x="Module",
       subtitle="Labels show uncapped absorption", y="Absorption (%)", caption="*Max bar height set at 200%")
cat("\n\n\\pagebreak\n")

 subset_hiv_melt = subset_hiv[start_date=="2019-01-01", .(abbrev_mod, label2, budget, expenditure)]
  subset_hiv_melt = melt(subset_hiv_melt, id.vars=c('abbrev_mod', 'label2'), variable.name='variable', value.name='amount')
  subset_hiv_melt[variable=="budget", label2:=""] #Don't display the expenditure amount on the budget bar. 
  
  #Fix budget and expenditure so they display nicely 
  subset_hiv_melt[variable=="budget", variable:="Budget"]
  subset_hiv_melt[variable=="expenditure", variable:="Expenditure"]
  
  ggplot(subset_hiv_melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label2), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Budget and expenditure for \nHIV and HIV/TB grants by module"), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")
  cat("\n\n\\pagebreak\n")
  
  cumulative = subset_hiv[, .(abbrev_mod, semester, budget, expenditure)]
  cumulative = melt(cumulative, id.vars=c('abbrev_mod', 'semester'))
  cumulative = dcast(cumulative, abbrev_mod+variable~semester)
  
  #Calculate cumulative budget and expenditure, and then absorption percentage. 
  cumulative[!is.na(`Semester 2`), cumulative:=sum(`Semester 1`, `Semester 2`, `Semester 3`), by=c('abbrev_mod', 'variable')]
  cumulative[!is.na(`Semester 1-2`), cumulative:=sum(`Semester 1-2`, `Semester 3`), by=c('abbrev_mod', 'variable')]
  stopifnot(nrow(cumulative[is.na(cumulative)])==0)
  
  #Subset this so it's easier to work with, and then graph. 
  cumulative_by_disease = cumulative[, .(abbrev_mod, variable, cumulative)]
  cumulative_by_disease = dcast(cumulative_by_disease, abbrev_mod~variable)
  cumulative_by_disease[, absorption:=(expenditure/budget)*100]
  cumulative_by_disease[, label1:=paste0(round(absorption, 1), "%")]
  cumulative_by_disease[absorption>200, absorption:=200] # Cap absorption at 200%
  cumulative_by_disease[, label2:=paste0(abbrev_mod)]
  
  ggplot(cumulative_by_disease, aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="indianred3") + 
  geom_text(hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Cumulative absorption for HIV, HIV/TB grants \nby module (Semesters 1-3)"), subtitle="Labels show uncapped absorption percentage", x="Module", y="Absorption (%)", caption="*Max bar height set at 200%")
  cat("\n\n\\pagebreak\n")
  
  melt = cumulative_by_disease[, .(abbrev_mod, budget, expenditure)]
  melt = melt(melt, id.vars=c('abbrev_mod'), variable.name='variable', value.name='amount')
  melt[variable=="budget", label:=""] #Don't display the expenditure amount on the budget bar. 
  melt[variable=="expenditure", label:=dollar(amount)]
  
  #Fix budget and expenditure so they display nicely 
  melt[variable=="budget", variable:="Budget"]
  melt[variable=="expenditure", variable:="Expenditure"]
  
  ggplot(melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_fill_manual(values=c('plum2', 'chocolate2')) + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Cumulative budget and expenditure for S1 2019"), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")

```

\newpage 
# Supporting data for pooled HIV/TB & HIV graphs
```{r supporting_data} 
kable(subset_hiv[, .(abbrev_mod, start_date, semester, budget, expenditure, absorption)], col.names=c('Module', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```
\newpage 

```{r by_module_combined2, results="asis"} 
subset_malaria = dt[grant_disease%in%c('malaria'), .(expenditure=sum(expenditure, na.rm=T), budget=sum(budget, na.rm=T)), by=c('semester', 'start_date', 'abbrev_mod')]
subset_malaria[, absorption:=(expenditure/budget)*100]
subset_malaria[, label1:=paste0(round(absorption, 1), "%")]
subset_malaria[absorption>200, absorption:=200] # Cap absorption at 200%
subset_malaria[, label2:=dollar(expenditure)]

ggplot(subset_malaria[start_date=="2019-01-01"], aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="darkorange1") + 
  geom_text(size=4, hjust=0) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Absorption by module, for malaria grants"), x="Module", y="Absorption (%)", caption="*Max bar height set at 200%", subtitle="Labels show uncapped absorption")
cat("\n\n\\pagebreak\n")

 subset_malaria_melt = subset_malaria[start_date=="2019-01-01", .(abbrev_mod, label2, budget, expenditure)]
  subset_malaria_melt = melt(subset_malaria_melt, id.vars=c('abbrev_mod', 'label2'), variable.name='variable', value.name='amount')
  subset_malaria_melt[variable=="budget", label2:=""] #Don't display the expenditure amount on the budget bar. 
  
  #Fix budget and expenditure so they display nicely 
  subset_malaria_melt[variable=="budget", variable:="Budget"]
  subset_malaria_melt[variable=="expenditure", variable:="Expenditure"]
  
  ggplot(subset_malaria_melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label2), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Budget and expenditure for \nmalaria grants by module"), x="Module", y="Amount ($)", fill="", subtitle="Labels show expenditure amounts")
  cat("\n\n\\pagebreak\n")
  
    
  cumulative = subset_malaria[, .(abbrev_mod, semester, budget, expenditure)]
  cumulative = melt(cumulative, id.vars=c('abbrev_mod', 'semester'))
  cumulative = dcast(cumulative, abbrev_mod+variable~semester)
  
  #Calculate cumulative budget and expenditure, and then absorption percentage. 
  cumulative[!is.na(`Semester 2`), cumulative:=sum(`Semester 1`, `Semester 2`, `Semester 3`, na.rm=T), by=c('abbrev_mod', 'variable')]
  cumulative[!is.na(`Semester 1-2`), cumulative:=sum(`Semester 1-2`, `Semester 3`, na.rm=T), by=c('abbrev_mod', 'variable')]
  stopifnot(nrow(cumulative[is.na(cumulative)])==0)
  
  #Subset this so it's easier to work with, and then graph. 
  cumulative_by_disease = cumulative[, .(abbrev_mod, variable, cumulative)]
  cumulative_by_disease = dcast(cumulative_by_disease, abbrev_mod~variable)
  cumulative_by_disease[, absorption:=(expenditure/budget)*100]
  cumulative_by_disease[, label1:=paste0(round(absorption, 1), "%")]
  cumulative_by_disease[absorption>200, absorption:=200] # Cap absorption at 200%
  cumulative_by_disease[, label2:=paste0(abbrev_mod)]
  
  ggplot(cumulative_by_disease, aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="indianred3") + 
  geom_text(hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Cumulative absorption for malaria grants \nby module (Semesters 1-3)"), subtitle="Labels show uncapped absorption percentage", x="Module", y="Absorption (%)", caption="*Max bar height set at 200%")
  cat("\n\n\\pagebreak\n")
  
    melt = cumulative_by_disease[, .(abbrev_mod, budget, expenditure)]
  melt = melt(melt, id.vars=c('abbrev_mod'), variable.name='variable', value.name='amount')
  melt[variable=="budget", label:=""] #Don't display the expenditure amount on the budget bar. 
  melt[variable=="expenditure", label:=dollar(amount)]
  
  #Fix budget and expenditure so they display nicely 
  melt[variable=="budget", variable:="Budget"]
  melt[variable=="expenditure", variable:="Expenditure"]
  
  ggplot(melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_fill_manual(values=c('plum2', 'chocolate2')) + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Cumulative budget and expenditure for S1 2019"), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")
```

\newpage 
# Supporting data for pooled malaria graphs
```{r supporting_data_mal} 
kable(subset_malaria[, .(abbrev_mod, start_date, semester, budget, expenditure, absorption)], col.names=c('Module', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```
\newpage 

```{r by_module_combined3, results="asis"} 
subset_tb = dt[grant_disease%in%c('tb'), .(expenditure=sum(expenditure, na.rm=T), budget=sum(budget, na.rm=T)), by=c('semester', 'start_date', 'abbrev_mod')]
subset_tb[, absorption:=(expenditure/budget)*100]
subset_tb[, label1:=paste0(round(absorption, 1), "%")]
subset_tb[absorption>200, absorption:=200] # Cap absorption at 200%
subset_tb[, label2:=dollar(expenditure)]

ggplot(subset_tb[start_date=="2019-01-01"], aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="slateblue2") + 
  geom_text(size=4, hjust=0) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Absorption by module, for TB grants"), x="Module", y="Absorption (%)", caption="*Max bar height set at 200%", subtitle="Labels show uncapped absorption")
cat("\n\n\\pagebreak\n")

 subset_tb_melt = subset_tb[start_date=="2019-01-01", .(abbrev_mod, label2, budget, expenditure)]
  subset_tb_melt = melt(subset_tb_melt, id.vars=c('abbrev_mod', 'label2'), variable.name='variable', value.name='amount')
  subset_tb_melt[variable=="budget", label2:=""] #Don't display the expenditure amount on the budget bar. 
  
  #Fix budget and expenditure so they display nicely 
  subset_tb_melt[variable=="budget", variable:="Budget"]
  subset_tb_melt[variable=="expenditure", variable:="Expenditure"]
  
  ggplot(subset_tb_melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label2), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  scale_y_continuous(labels = scales::dollar, limits=c(0, 6000000)) + 
  labs(title=paste0("Budget and expenditure for TB grants by module"), x="Module", y="Amount ($)", fill="", subtitle="Labels show expenditure amounts")
  cat("\n\n\\pagebreak\n")
    
  cumulative = subset_tb[, .(abbrev_mod, semester, budget, expenditure)]
  cumulative = melt(cumulative, id.vars=c('abbrev_mod', 'semester'))
  cumulative = dcast(cumulative, abbrev_mod+variable~semester)
  
  #Calculate cumulative budget and expenditure, and then absorption percentage. 
  cumulative[, cumulative:=sum(`Semester 1`, `Semester 2`, `Semester 3`, na.rm=T), by=c('abbrev_mod', 'variable')]
  stopifnot(nrow(cumulative[is.na(cumulative)])==0)
  
  #Subset this so it's easier to work with, and then graph. 
  cumulative_by_disease = cumulative[, .(abbrev_mod, variable, cumulative)]
  cumulative_by_disease = dcast(cumulative_by_disease, abbrev_mod~variable)
  cumulative_by_disease[, absorption:=(expenditure/budget)*100]
  cumulative_by_disease[, label1:=paste0(round(absorption, 1), "%")]
  cumulative_by_disease[absorption>200, absorption:=200] # Cap absorption at 200%
  cumulative_by_disease[, label2:=paste0(abbrev_mod)]
  
  ggplot(cumulative_by_disease, aes(x=abbrev_mod, y=absorption, label=label1))+ 
  geom_bar(stat="identity", fill="indianred3") + 
  geom_text(hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Cumulative absorption for TB grants \nby module (Semesters 1-3)"), subtitle="Labels show uncapped absorption percentage", x="Module", y="Absorption (%)", caption="*Max bar height set at 200%")
  cat("\n\n\\pagebreak\n")
  
    melt = cumulative_by_disease[, .(abbrev_mod, budget, expenditure)]
  melt = melt(melt, id.vars=c('abbrev_mod'), variable.name='variable', value.name='amount')
  melt[variable=="budget", label:=""] #Don't display the expenditure amount on the budget bar. 
  melt[variable=="expenditure", label:=dollar(amount)]
  
  #Fix budget and expenditure so they display nicely 
  melt[variable=="budget", variable:="Budget"]
  melt[variable=="expenditure", variable:="Expenditure"]
  
  ggplot(melt, aes(x=abbrev_mod, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  coord_flip() + 
  scale_fill_manual(values=c('plum2', 'chocolate2')) + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Cumulative budget and expenditure for S1 2019"), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")
```

\newpage 
# Supporting data for pooled TB graphs
```{r supporting_data_tb} 
kable(subset_tb[, .(abbrev_mod, start_date, semester, budget, expenditure, absorption)], col.names=c('Module', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```
\newpage 

```{r by_module_counts, include=FALSE}
labels = by_module[start_date=="2019-01-01", .(num_grants=.N), by=c('abbrev_mod')]
subset = merge(subset, labels, by='abbrev_mod')
subset[, num_grants:=as.character(num_grants)] #Turn this into a label. 

ggplot(subset[start_date=="2019-01-01"], aes(x=abbrev_mod, y=absorption, label=num_grants))+ 
  geom_bar(stat="identity", fill="mediumorchid1") + 
  geom_text(hjust=-0.5, size=6) + 
  theme_bw(base_size=16) + 
  coord_flip() + e
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title=paste0("Absorption by module, all grants"), x="Module", y="Absorption (%)", caption="*Max bar height set at 200%")
```


\newpage 

# RSSH absorption across grants 

\newpage
```{r rssh}
rssh_mods = c('Community systems', 'Financial systems', "Info systems & M&E", "HR & health workers", "Service delivery", "Nat. health strategies")

rssh_by_int = dt[abbrev_mod%in%rssh_mods & start_date=="2019-01-01", .(budget=sum(budget, na.rm=T), expenditure=sum(expenditure, na.rm=T)), by=c('abbrev_mod', 'abbrev_int', 'start_date')]
rssh_by_int[, absorption:=(expenditure/budget)*100]
rssh_by_int[, label1:=paste0(round(absorption, 1), "%")]
rssh_by_int[absorption>200, absorption:=200] # Cap absorption at 200%

#Shorten a few intervention names. 
rssh_by_int[abbrev_int=="Supportive policy and programmatic environment", abbrev_int:="Supportive policy environment"]

#Sort the graph by module 
rssh_by_int[, module_num:=.GRP, by='abbrev_mod']
rssh_by_int[, int_num:=.GRP, by=c('abbrev_mod', 'abbrev_int')]
rssh_by_int[, rssh_factor:=paste0(module_num, int_num)]

rssh_by_int$int_factor = factor(rssh_by_int$rssh_factor, unique(rssh_by_int$rssh_factor), unique(rssh_by_int$abbrev_int))

ggplot(rssh_by_int, aes(x=int_factor, y=absorption, fill=abbrev_mod, label=label1))+ 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(hjust=-0.5, size=4) + 
  theme_bw(base_size=14) + 
  scale_fill_brewer(palette="Set2") + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title="2019 RSSH absorption by intervention, all grants", subtitle="Interventions grouped by module", x="Intervention", y="Absorption (%)",   fill="Module", caption="*Max bar height set at 200%")
```

```{r cumulative_rssh}

#Show cumulative RSSH absorption
  cumulative_rssh = dt[abbrev_mod%in%rssh_mods, .(budget=sum(budget, na.rm=T), expenditure=sum(expenditure, na.rm=T)), by=c('grant', 'abbrev_mod', 'abbrev_int', 'semester')]
  cumulative_rssh = melt(cumulative_rssh, id.vars=c('grant', 'abbrev_mod', 'abbrev_int', 'semester'))
  cumulative_rssh = dcast(cumulative_rssh, grant+abbrev_mod+abbrev_int+variable~semester) #Need to collapse to the grant-level first, so you can get cumulative variable calculated right for different PUDR reporting schemes. Then collapse to intervention level. 
  
  #Calculate cumulative budget and expenditure, and then absorption percentage. 
  cumulative_rssh[!is.na(`Semester 2`), cumulative:=sum(`Semester 1`, `Semester 2`, `Semester 3`, na.rm=T), by=c('abbrev_mod', 'abbrev_int', 'variable', 'grant')]
  cumulative_rssh[!is.na(`Semester 1-2`), cumulative:=sum(`Semester 1-2`, `Semester 3`, na.rm=T), by=c('abbrev_mod', 'abbrev_int', 'variable', 'grant')]
  cumulative_rssh[grant=="UGA-M-MoFPED", cumulative:=sum(`Semester 1`, `Semester 2`, `Semester 3`, na.rm=T), by=c('abbrev_mod', 'abbrev_int', 'variable', 'grant')] #This grant had some new interventions added in Semester 3. 
  stopifnot(nrow(cumulative_rssh[is.na(cumulative)])==0)
  
  #Subset this so it's easier to work with, and then graph. 
  cumulative_rssh = cumulative_rssh[, .(cumulative=sum(cumulative, na.rm=T)), by=c('abbrev_mod', 'abbrev_int', 'variable')] #Collapse grant out
  cumulative_rssh = dcast(cumulative_rssh, abbrev_mod+abbrev_int~variable)
  cumulative_rssh[, absorption:=(expenditure/budget)*100]
  cumulative_rssh[, label1:=paste0(round(absorption, 1), "%")]
  cumulative_rssh[absorption>200, absorption:=200] # Cap absorption at 200%
  cumulative_rssh[, label2:=paste0(abbrev_mod)]
  
  #Shorten a few intervention names. 
  cumulative_rssh[abbrev_int=="Supportive policy and programmatic environment", abbrev_int:="Supportive policy environment"]
  
  #Sort the graph by module 
  cumulative_rssh[, module_num:=.GRP, by='abbrev_mod']
  cumulative_rssh[, int_num:=.GRP, by=c('abbrev_mod', 'abbrev_int')]
  cumulative_rssh[, rssh_factor:=paste0(module_num, int_num)]
  
  cumulative_rssh$int_factor = factor(cumulative_rssh$rssh_factor, unique(cumulative_rssh$rssh_factor), unique(cumulative_rssh$abbrev_int))
  
  ggplot(cumulative_rssh, aes(x=int_factor, y=absorption, fill=abbrev_mod, label=label1))+ 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(hjust=-0.5, size=4) + 
  theme_bw(base_size=14) + 
  scale_fill_brewer(palette="Set2") + 
  coord_flip() + 
  scale_y_continuous(limits=c(0, 220)) + 
  labs(title="Cumulative RSSH absorption by intervention, all grants", subtitle="(Semesters 1-3) \nInterventions grouped by module", x="Intervention", y="Absorption (%)",   fill="Module", caption="*Max bar height set at 200%")

melt = cumulative_rssh[, .(abbrev_mod, abbrev_int, int_factor, budget, expenditure)]
  melt = melt(melt, id.vars=c('abbrev_mod', 'abbrev_int', 'int_factor'), variable.name='variable', value.name='amount')
  melt[variable=="budget", label:=""] #Don't display the expenditure amount on the budget bar. 
  melt[variable=="expenditure", label:=dollar(amount)]
  
  #Fix budget and expenditure so they display nicely 
  melt[variable=="budget", variable:="Budget"]
  melt[variable=="expenditure", variable:="Expenditure"]
  
  ggplot(melt, aes(x=int_factor, y=amount, fill=variable))+ 
  geom_bar(stat="identity", position="identity") + 
  geom_text(aes(label=label), hjust=0, size=4) + 
  theme_bw(base_size=16) + 
  scale_fill_manual(values=c('plum2', 'chocolate2')) + 
  coord_flip() + 
  scale_y_continuous(labels = scales::dollar) + 
  labs(title=paste0("Cumulative RSSH budget and expenditure"), x="Module", y="Amount ($)", fill="", 
       subtitle="Labels show expenditure amounts")
```


# Supporting data for RSSH graphs (1/2)
```{r supporting_data_rssh1} 
 kable_data = dt[abbrev_mod%in%rssh_mods, .(budget=sum(budget, na.rm=T), expenditure=sum(expenditure, na.rm=T)), by=c('abbrev_mod', 'abbrev_int', 'start_date', 'semester')]
kable_data[, absorption:=(expenditure/budget)*100]

kable(kable_data[1:30, .(abbrev_mod, abbrev_int, start_date, semester, budget, expenditure, absorption)][order(abbrev_mod, abbrev_int, start_date, semester)], col.names=c('Module', 'Intervention', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```

# Supporting data for RSSH graphs (2/2)
```{r supporting_data_rssh2} 
stopifnot(nrow(kable_data)==42)
kable(kable_data[31:42, .(abbrev_mod, abbrev_int, start_date, semester, budget, expenditure, absorption)][order(abbrev_mod, abbrev_int, start_date, semester)], col.names=c('Module', 'Intervention', 'Date', 'Semester', 'Budget', 'Expenditure', 'Absorption'), align="l") %>%
  kable_styling(font_size=8)
```
