---
title: "Guatemala Performance Indicators"
author: "Francisco Rios Casas"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(data.table)
library(ggplot2)
library(knitr)
library(dplyr)
library(kableExtra)
library(scales)

# set parameters of data to be analyzed
country = "gtm"

main_indicators = c('impact_outcome_indicators_main', 'coverage_indicators_main')
end = "2019-06-30"


#recent_pudrs = c("GTM-H-HIVOS_Progress Report_31Dec2018_RV_SO_YAV_120219_LFA REVIEW Reviewed FS.xlsx",
#                 "GTM-M-MSPAS_PU_FINAL_fuentes_indi_corregidos.xlsx",
#                 "GTM-T-MSPAS_Progress Report_LFA18Mar19.xlsx")

LFA_verified_grants = c("GTM-H-HIVOS_Progress Report_31Dec2018_RV_SO_YAV_120219_LFA REVIEW Reviewed FS.xlsx",
                        "GTM-T-MSPAS_Progress Report_LFA18Mar19.xlsx") # file names of most-recent LFA-verified grants

# load data
data <- readRDS("J:/Project/Evaluation/GF/outcome_measurement/multi_country/performance_indicators/pudr_indicator_extraction/cleaned_data/kpi_data_for_analyses2.RDS")

# subset as appropriate
DT <- data
DT = DT[loc_name==country & pudr_sheet %in% main_indicators & end_date_programmatic==end]
```


## 1. Introduction  

This document highlights the results of Performance Indicators in Guatemala. Data for indicators all came from the latest available Progress Update and Disbursement Request forms (PUDRs) indicated below:  


```{r toc}
recent_pudrs = unique(DT[,.(file_name, start_date_programmatic, end_date_programmatic)])
kable(recent_pudrs, col.names=c("File", "Start of Reporting Period", "End of Reporting Period")) %>%
  column_spec(c(2,3),width = "10em")
```

**There are two grants reporting for the most recent time period S1 2019.**


## 2. Grant Results based on most recent PUDRs  


The "Achievement Ratio" is calculated in the PUDR to understand how close to the original target the Outcome Indicator result is. A value of 1.0 or greater means that the Indicator's goal has been met or exceeded.


```{r plot_results}
# plots grant performance in a loop for all grants in country
DT[ihme_result_achievement_ratio>2.0, ihme_result_achievement_ratio:=2.0]
grants = unique(DT$grant)
plots = list()
i=1
for(g in grants) {
plots[[i]] = ggplot(DT[grant==g], aes(x=brief_description_code, y=ihme_result_achievement_ratio, col=target_met)) +
  geom_point() +
  labs(col="Target Met", title="", y="Achievement Ratio", x="Indicator", caption=paste0("Source: ", unique(DT[grant==g]$file_name))) +
  geom_hline(yintercept = 1) +
  theme_bw()+
  coord_flip()+
  ylim(0,1.5)+
  theme(legend.position = "bottom")+
  theme_bw()
i=i+1
}
```

### TB Grant

```{r}
plots[[1]]
```

**Indicators related to TB and TB/HIV mortality, TB incidence, and case notifications are on-track. The indicator furthest from its expected target concerns children in contact with TB patients who began isonizid preventive therapy. Additionally, it seems that in general, indicators that concern MDR-TB are also below the expected targets.**


### HIV Grants
```{r}
plots[[2]]
```

**The indicators furthest from their targets relate to linkage to care for people newly diagnosed with HIV and HIV testing and prevention programs for Sex workers and MSM.**

```{r}
# subset as appropriate
DT2 <- data
country = "gtm"

DT2 = DT2[loc_name==country & pudr_sheet %in% main_indicators & start_date_programmatic=="2018-07-01" & end_date_programmatic=="2018-12-31"]

DT2[ihme_result_achievement_ratio>2.0, ihme_result_achievement_ratio:=2.0]
grants2 = unique(DT2$grant)
bplots = list()
i=1
for(g in grants2) {
bplots[[i]] = ggplot(DT2[grant==g], aes(x=brief_description_code, y=ihme_result_achievement_ratio, col=target_met)) +
  geom_point() +
  labs(col="Target Met", title="", y="Achievement Ratio", x="Indicator", caption=paste0("Source: ", unique(DT2[grant==g]$file_name))) +
  geom_hline(yintercept = 1) +
  theme_bw()+
  coord_flip()+
  ylim(0,1.5)+
  theme(legend.position = "bottom")+
  theme_bw()
i=i+1
}
```

```{r}
bplots[[3]]
```

**Results available for S2 2018.**

### Malaria Grant

```{r}
bplots[[1]]
```

**Majority of indicators are close even if they are not meeting their targets. Stock-out of key commodities is the indicator that is furthest from target.**

**Results available for S2 2018.**