---
title: "Uganda Performance Indicators"
author: "Francisco Rios Casas"
date: "10/29/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(data.table)
library(ggplot2)
library(knitr)
library(dplyr)
library(kableExtra)
library(scales)

# set parameters of data to be analyzed
country = "uga"
main_indicators = c('impact_outcome_indicators_main', 'coverage_indicators_main')
recent_pudrs = c("Final LFA reviewed UGA-M-MoFPED PUDR.xlsx",
                 "UGA-M-TASO PUDR Jul-Dec18.xlsx",
                 "UGA-H-MoFPED Progress Update Report Jan-Jun 2019.xlsx",
                 "LFA Reviewed UGA-C-TASO PE 31Dec18.xlsx",
                 "LFA Reviewed UGA-T-MOFPED PE 31 Dec 2019 (10 May 2019).xlsx"
                 )
LFA_verified_grants = c("Final LFA reviewed UGA-M-MoFPED PUDR.xlsx",
                        "Final LFA reviewed UGA-M-TASO PUDR 2 Oct 2018.xlsx",
                        "LFA Reviewed UGA-H-MoFPED PUDR PE 31 Dec 18  (10 May 19).xlsx",
                        "LFA Reviewed UGA-C-TASO PE 31Dec18.xlsx",
                        "LFA Reviewed UGA-T-MOFPED PE 31 Dec 2019 (10 May 2019).xlsx") # file names of most-recent LFA-verified grants

# load data
data <- readRDS("J:/Project/Evaluation/GF/outcome_measurement/multi_country/performance_indicators/pudr_indicator_extraction/cleaned_data/kpi_data_for_analyses.RDS")

# subset as appropriate
DT <- data
DT = DT[loc_name==country]
DT = DT[pudr_sheet %in% main_indicators]
DT = DT[file_name %in% recent_pudrs]
```


## 1. Introduction  

This document highlights the data completeness, data sources, and changes over time for Outcome Indicators in Uganda. Data for indicators all came from the latest available Progress Update and Disbursement Request forms (PUDRs) indicated below:  


```{r toc}
kable(recent_pudrs, col.names = c("Most Recent PUDRs available")) %>%
  kable_styling(bootstrap_options = "striped", position = "center")
```

## 2. Data Completeness  

For each Outcome Indicator, the grant Principal Recipient (PR) *usually* must establish an indicator "Baseline Value". This value--when available--can serve as a benchmark for progress reported in subsequent time periods. 

```{r completeness}
# Emily's code on {PUDR} Completeness

# Data
# What percentage of targets do they report on?
# Convert NAs to "Unknown"


# Convert NAs to "Unknown"
DT$plot_baseline <-NA
DT$plot_baseline[is.na(DT$baseline_value)] <- "Baseline Not Reported"
DT$plot_baseline[!is.na(DT$baseline_value)] <- "Baseline Value Reported"

ggplot(DT, aes(x=grant, fill=plot_baseline))+
  geom_bar(position = "fill")+
  labs(title="Proportion of PUDRs that report on baseline values", y="")+
  theme_bw(base_size = 8)+
  theme(legend.title=element_blank())+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
```


**Baseline values are available for between 25% to 50% of indicators in the PUDRs.**


In order to calculate the achievement of an Outcome Indicator for a given grant, it is necessary to have information on (1) the target that was originally set and (2) the result.

```{r completeness2}
# create completeness indicator for baseline seperately from target and result

ggplot(DT, aes(x = grant, fill = completeness_rating)) +
  geom_bar(position = "fill")+
  labs(title="Completeness of Indicators according to grants", y="")+
  theme_bw(base_size = 8)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  theme(legend.title=element_blank())
```

**There is a high amount of data completeness across the Uganda grants. The majority of grants reported both a result and a target to measure progress.**  
 
 
## 3. Data Sources  

"Data Sources" are listed in the PUDR to indicate where information reported to the global fund came from. Examples of data sources include reports from the disease-specific National Programs, surveys such as the Demographic Health Survey (DHS), and Health Management Information systems such as DHIS2.  

In some cases, the source of information used to establish the target of a goal is different from the source used to provide updates on progress. **In Uganda, usually the data sources changed very little between what was used to establish the goal and what was used to provide an update.**  


```{r sources}
# create indicator variable of when result and target source differ
# subset data for checking changing indicators
table3 <-DT[,list(indicators=length(indicator_code), total_different=sum(sources_different, na.rm = TRUE)), by=c('grant')]

kable(table3, col.names = c("Grant", "Indicators", "Different Sources between Baseline and Result")) %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em") %>%
  kable_styling(bootstrap_options = "striped", position = "center")

```

  
**When data sources did change as in the case of the UGA-M-TASO and UGA-T-MoFPED grants, the original data sources were mostly survey data or annual reports while results were obtained through the country's Health Management Information System (HMIS).**   


```{r sources_appendix1, include=FALSE}
# subset data for checking changing indicators
table3b <-DT[grant=="UGA-T-MoFPED",.(indicator_code, brief_description_code, baseline_source, pr_result_source)]

kable(table3b, col.names = c("Indicator", "Description", "Baseline Source", "Result source"))


```

```{r sources_appendix2, include=FALSE}
table3c <-DT[grant=="COD-H-MOH",.(indicator_code, brief_description_code, baseline_source_code, pr_result_source_code, sources_different)]

kable(table3c, col.names = c("Indicator", "Description", "Baseline Source", "Result source", "indicx"))

```
  

## 4. Grant Results based on most recent PUDRs  


The "Achievement Ratio" is calculated in the PUDR to understand how close to the original target the Outcome Indicator result is. A value of 1.0 or greater means that the Indicator's goal has been met or exceeded.


```{r plot_results}
# plots grant performance in a loop for all grants in country
grants = unique(DT$grant)
plots = list()
i=1
for(g in grants) {
plots[[i]] = ggplot(DT[grant==g], aes(x=brief_description_code, y=ihme_result_achievement_ratio)) +
  geom_point() +
  labs(title="", y="Achievement Ratio", x="Indicator", caption=paste0("Source: ", unique(DT[grant==g]$file_name))) +
  geom_hline(yintercept = 1) +
  theme_bw()+
  coord_flip()+
  ylim(0,1.5)+
  theme(legend.position = "bottom")+
  theme_bw()
i=i+1
}
```

### 4.1 Malaria Grants

```{r}
plots[[1]]
```

**A drop in reported malaria cases has meant that the indicator for this value (I-1(M)) has actually far exceeded its targets (in this case an achievement ratio below 1 indicates that the goal has been met or exceeded.) The indicators further from their original goals are the number of long-lasting insecticidal treated nets (LLIN) distributed and the proportion of women attending antenatal care.**

```{r}
plots[[2]]
```

**The lowest-achieving indicators in the UGA-M-TASO grant (not already reported in the previous malaria grant above) are related to antimalarial treatment in private sector sites and the timely reporting of HMIS-reporting sites.**


### 4.2 HIV
```{r}
plots[[3]]
```
**More than half of UGA-H-MoFPED indicators are meeting or exceeding their targets. The indicator furthest from its target relates to TB preventive therapy for people living with HIV.**

### 4.3 Combined Grant
```{r}
plots[[4]]
```
**The main TB indicators in the UGA-C-TASO grant are exceeding or close to meeting their targets, but those relating to specific sub-populations are further from their goals. These include: adolescent girls and young women (AGYW), men who have sex with men (MSM), and sex workers (SW).** 

### 4.4 TB Grant
```{r}
plots[[5]]
```

**Most of the indicators reported for the UGA-T-MoFPED grant are close to their targets. Treatment for MDR-TB and Rifampicin drug-susceptibility testing were the furthest from their goals, but not by a large amount.**

## 5. Agreement between PR, LFA

After each PUDR is submitted by the PR, the LFA will verify the results reported on the Indicators. In some cases they may revise the original values provided by the PR.

The most recent LFA-verified PUDRs for this analysis include the following:


```{r}
kable(LFA_verified_grants, col.names = c("LFA Verified PUDR Files")) %>%
  kable_styling(bootstrap_options = "striped", position = "center")
```



```{r lfa_agreement}

data.agreement <- data[file_name%in%LFA_verified_grants & lfa_result_value<=200 & pr_result_value <=200] # scale to make sure outliers don't disrupt the rest of the graph

ggplot(data.agreement, aes(x=pr_result_value, y=lfa_result_value, color=grant))+
  geom_point()+
  labs(x="PR Reported Value", y="LFA Verified Values", title = paste0("Agreement between PR and LFA in grants"))+
  geom_abline(slope = 1, intercept = 0)+
  theme_bw()+
  scale_colour_discrete(name="Grants")
  #+ geom_text(data=data.agreement[pr_result_value>60 & lfa_result_value<50], aes(label=indicator_code), hjust="right", check_overlap = TRUE)

```


**In general, there were few instances where the GF LFA had to change the results reported by the PRs. In the few instances where it did happen, the LFA usually would increase the values that the PR reported.**


## 6. Results Over Time  

The tables below highlights how indicators available in the three most recent PUDRs have changed over time for each of the grants.
```{r}
# subset data

# subset parameters
variables_to_plot = c("ihme_result_achievement_ratio")

# subset to most recent pudrs and main_indicators
data2 <- data[loc_name==country & pudr_sheet%in%main_indicators]

# re-shape the data
data2 <- melt(data2, id.vars = c("indicator_code", "grant", "brief_description_code", "end_date_programmatic", "file_name", "type", "pudr_sheet"))
data2 <- data2[variable %in% variables_to_plot]

# change format
data2$value <- as.numeric(data2$value)
data2 <- data2[!is.na(data2$value),]

# subset data to only those indicators with three time points per grant
keep <- data2[,.N,by=indicator_code] #keep indicators with 3 time points
indicators <- keep[N==3,indicator_code] # create vector of indicators with 3 time point stop
data2 <- data2[indicator_code %in% indicators]

# plot the data
# write a loop to make plots for each grant
grants = unique(data2$grant)
dplots = list()
i=1
for(g in grants) {
dplots[[i]] = ggplot(data2[grant==g], aes(x=end_date_programmatic, y=value, group=brief_description_code, color=brief_description_code))+
  geom_point()+
  geom_line()+
  theme_bw()+
  scale_x_date(date_breaks = "6 month", 
                 labels=date_format("%b-%Y"))+
   labs(x="Date of PUDR", y="Achievement Ratio", title=paste0(g,"\n", "Performance Over Time"))+
  scale_colour_discrete(name="")+
  geom_hline(yintercept=1, linetype="dashed", color = "gray")+
  ylim(0, 1.5)+
  theme(legend.position="bottom", legend.text=element_text(size=6))+
  guides(col=guide_legend(ncol=1, keywidth = ))

i=i+1
}
```

```{r}
dplots[[1]]
```
**In general, indicator results in the UGA-H-MoFPED grant have remained very stable over the past few semesters. TB-preventive treatment for people living with HIV (PLHIV) has continuously been the lowest-achieving indicator in the past few reporting time periods.**


```{r}
dplots[[3]]
```
**The only indicator for the UGA-M-MoFPED grant with multiple time-points is number of LLIN nets distributed. This indicator has not reached its goals within the past few semesters.**

```{r, include=FALSE}
# target-setting over time

# subset parameters
trg_variables_to_plot = c("target_value")

# subset to most recent pudrs and main_indicators
data3 <- data[loc_name==country & pudr_sheet%in%main_indicators]

# re-shape the data
data3 <- melt(data3, id.vars = c("indicator_code", "grant", "brief_description_code", "end_date_programmatic", "file_name", "type", "pudr_sheet"))
data3 <- data3[variable %in% trg_variables_to_plot]

# change format
data3$value <- as.numeric(data3$value)
data3 <- data3[!is.na(data3$value),]

# subset data to only those indicators with three time points per grant
keeps <- data3[,.N,by=indicator_code] #keep indicators with 3 time points
trg_indicators <- keeps[N==3,indicator_code] # create vector of indicators with 3 time point stop
data3 <- data3[indicator_code %in% trg_indicators]

grants = unique(data3$grant)

cplots = list()
i=1
for(g in grants) {
cplots[[i]] = ggplot(data3[grant==g], aes(x=end_date_programmatic, y=value, group=brief_description_code, col=brief_description_code))+
  geom_point()+
  geom_line()+
  theme_bw()+
  facet_wrap(~variable)+
  scale_x_date(date_breaks = "6 month", 
                 labels=date_format("%b-%Y"))+
   labs(x="Date of PUDR", y="Target Value", title=paste0(g, "\n", "Target-setting Over Time"), caption = paste0("Data source: ", unique(data3$file_name)))+
  scale_colour_discrete(name="")+
  theme(legend.position="bottom")+
  guides(col=guide_legend(ncol=1))
  
i=i+1
}

```
