---
title: "PNLS PMTCT Quality Checks"
author: "Emily Linebarger"
date: "June 21, 2019"
output: html_document
---

```{r setup, echo=FALSE, warning=FALSE}
#Set up R
rm(list=ls())
library(data.table)
library(ggplot2)
library(knitr)
# --------------------

# shell script for working on the cluster
# sh /share/singularity-images/rstudio/shells/rstudio_qsub_script.sh -p 1247 -s 2 

# --------------------
# set working directories

# detect if operating on windows or on the cluster 
j = ifelse(Sys.info()[1]=='Windows', 'J:', '/home/j')

# set the directory for input and output
dir = paste0(j, '/Project/Evaluation/GF/outcome_measurement/cod/dhis_data/')
setwd(dir)

dt = readRDS(paste0(dir, 'prepped/pnls_final/pnls_pmtct.rds'))
dt[, year:=year(date)]

dt[, element_no:=.GRP, by='element_eng'] #Make a new numerical element identifier for new English names. 

#----------------------------------------------------------
# Define some useful functions for checks below 
#----------------------------------------------------------
#Given 2 element IDs, does the first one always have less than the second for date and org unit ID? 
subset_check = function(element1, element2){
  subset1 = dt[element_no==element1, .(value1=sum(value, na.rm=T)), by=c('org_unit_id', 'date')]
  subset2 = dt[element_no==element2, .(value2=sum(value, na.rm=T)), by=c('org_unit_id', 'date')]
  subset = merge(subset1, subset2, by=c('org_unit_id', 'date'), all=T)
  subset[, diff:=value2-value1]
  
  if (nrow(subset[!is.na(diff) & diff<0])){
    return(FALSE)
  }
  return(TRUE)
} 

#--------------------------------------------------------
# Variables of interest: 
# 1. Counseled, tested, and HIV+ for pregnant women
# 2. Tested and HIV+ for infants
#--------------------------------------------------------

```

## HIV Testing for pregnant and lactating Women 

```{r plw testing checks, echo=FALSE, warning=FALSE}
#---------------------------------------------------------------------
# Pregnant and lactating women variable overlap analysis
#---------------------------------------------------------------------
#What are the unique element IDs for pregnant and lactating women? 
plw = unique(dt[subpop=="plw", .(element_eng, element_no)])

#Generate a variable that represents the total number of PLW tested. 
tested_vars = plw[grep("tested", tolower(element_eng)), unique(element_no)]
kable(plw[element_no%in%tested_vars, unique(element_eng)])

#Do a visual check to make sure you haven't missed any 
plw[!element_no%in%tested_vars, .(element_eng)] #All are included EKL 6/21

#Compare the two variables that are included. 
plw_values = dt[element_no%in%tested_vars & subpop=="plw", .(value=sum(value, na.rm=T)), by=c('date', 'element_eng')]
plot = ggplot(plw_values, aes(x=date, y=value, color=element_eng))+
  geom_line() + geom_point() + 
  theme_bw() + 
  labs(title="Comparing overlap in PLW testing variables", x="Date", y="Value, summed by month")
plot

```

It looks like "Tested HIV+ in the delivery room and were put on ART after delivery" is very low; what are the unique values?
```{r testing2, echo=FALSE}
kable(unique(dt[element_no==51, .(value)]))
```
The highest number here is 21. 

*DECISION:*
Of the two variables "Tested" and "Tested HIV+ in the delivery room and put on ART after delivery", the number of cases of the second variable is so low that it should be sufficient to use the "Testing" variable only to analyze testing. 

## HIV Counseling for pregnant and lactating women
```{r plw Counseled checks, echo=FALSE, warning=FALSE}
#Generate a variable that represents the total number of PLW tested. 
counseled_vars = plw[grep("counseled", tolower(element_eng)), unique(element_no)]
kable(unique(plw[element_no%in%counseled_vars, .(element_eng, element_no)]))

```
There is only one 'Counseled' variable in this dataset. 


```{r HIV1, echo=FALSE, warning=FALSE}
#Generate a variable that represents the total number of PLW that are HIV+. 

positive_vars = plw[grep("HIV\\+", element_eng), unique(element_no)]
kable(unique(plw[element_no%in%positive_vars, .(element_eng)]))
final_positive_vars = copy(positive_vars)
```

These are the variables that aren't included in the list above - this split is correct EKL 6/21
```{r HIV2, echo=FALSE}
#Do a visual check to make sure you haven't missed any 

plw[!element_no%in%positive_vars, .(element_eng)] #All are included EKL 6/21

```

What do all of these variables look like? 
```{r hiv3, echo=FALSE}
x = dt[element_no %in% positive_vars]
x = x[ ,.(value=sum(value)), by=.(date, element_eng)]
plot = ggplot(x, aes(x=date, y=value, color=element_eng)) +
  geom_point() +
  geom_line()
plot 
```
```{r reporting, echo=FALSE}
#-------------------------------------------
#Do a reporting completeness graph 
#-------------------------------------------
FacsReportingByDate = dt[, .(facilities=length(unique(org_unit_id))), by=date]
FacsReportingDrugByDate = dt[subpop=="plw" & element_no%in%positive_vars, .(facs_reporting_on_drug=as.numeric(length(unique(org_unit_id)))), by=date]

#Did they report for the specific drug you're targeting?
report = merge(FacsReportingByDate, FacsReportingDrugByDate, by='date', all.x=TRUE)

report[is.na(facs_reporting_on_drug), facs_reporting_on_drug:=0]
report[ , drug_reporting_ratio:=100*(facs_reporting_on_drug/facilities)]

# convert to numerics and round
report[, drug_reporting_ratio:=round(drug_reporting_ratio, 1)]

plot = ggplot(report, aes(x=date, y=drug_reporting_ratio)) +
  geom_point(size=0.8) +
  geom_line() +
  geom_line() +
  theme_bw() +
  # facet_wrap(~indicator) +
  scale_color_manual(values = brewer.pal(4, 'RdYlBu')) +
  labs(x='Date', y='Number of health facilities', title='Number of facilities reporting on HIV+ indicators')
plot
#Make a series of graphs to explore the overlap between these variables. 
#----------------------------------------------------------------------------
# Graph all HIV+ variables for the month with the highest reporting. 
#----------------------------------------------------------------------------
positive_values = dt[element_no%in%positive_vars & subpop=="plw", .(value=sum(value, na.rm=T)), by=c('date', 'element_eng')]
positive_values = positive_values[order(date, -value)]
kable(positive_values[date=="2017-07-01"]) #Visual check before graph

plot = ggplot(positive_values[date=="2017-07-01"], aes(x=date, y=value, fill=element_eng))+
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  labs(title="Comparing overlap in PLW HIV+ variables for the month of July, 2017\n(selected for high reporting completeness)", x="Date", y="Value, summed by month")
plot


```


```{r other analyses, include=FALSE}
#Is "HIV+, receiving ART and gave birth in the maternity ward" always a subset of "HIV+ and gave birth in the maternity ward?"
  delivery2 = dt[element_no%in%c('H0skpxsq2U3', "CyZ9wMA7iSi"), .(value=sum(value, na.rm=T)), by=c('date', 'element_eng')]
  ggplot(delivery2, aes(x=date, y=value, fill=element_eng))+
    geom_bar(stat="identity", position="dodge") + 
    theme_bw() + 
    labs(title="Comparing overlap in PLW delivery/facility treatment variables", x="Date", y="Value, summed by month")
  
  #Also do through calculation 
  subset_check('H0skpxsq2U3', "CyZ9wMA7iSi")

  #DECISION - Unsure. there are some cases where you don't have full reporting overlap. 
  
  #Is HIV+ received in the facility during the month, and already receiving ART a subset of HIV+ received inthe facility during the month? 
  delivery3 = dt[element_no%in%c('iKekGMwvtVy', "NIC3bM4fv4v"), .(value=sum(value, na.rm=T)), by=c('date', 'element_eng')]
  ggplot(delivery3, aes(x=date, y=value, fill=element_eng))+
    geom_bar(stat="identity", position="dodge") + 
    theme_bw() + 
    labs(title="Comparing overlap in PLW delivery/facility treatment variables", x="Date", y="Value, summed by month")
  
  #Subset calculation 
  subset_check('iKekGMwvtVy', "NIC3bM4fv4v")
  #DECISION - Also seems likely. 
  final_positive_vars = final_positive_vars[final_positive_vars!="iKekGMwvtVy"]
  
  #Is HIV+ and received in the facility during the month always a subset of "Knew their HIV status before arriving at the facility for ART?"
  subset_check("NIC3bM4fv4v", "urDEeghmDdM")
  
  #2. Received ART 
  positive_subpops = dt[element_no%in%positive_vars & subpop=="plw", .(value=sum(value, na.rm=T)), by=c('date', 'element_eng', 'maternity', 'sex', 'case', 'age')]

```
```{r early_infant_checks, echo=FALSE, warning=FALSE}

#---------------------------------------------------------------------
# Infant variable overlap analysis
#---------------------------------------------------------------------
unique(dt[subpop=="exposed_infant", .(element_eng, element_no)])

```