---
title: "Research Team 3.15.19"
author: "Emily Linebarger"
date: "March 13, 2019"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
dt = readRDS(paste0(fgh_raw, "ghe_actuals_malaria.rds"))

drc = dt[ihme_loc_id == 'COD']
drc_ghe = drc[value_code == 'fs_malaria_domestic_public']
```

## Exploratory Analysis of GHE spending on malaria from FGH team 

Plot of government health expenditure on malaria 

```{r, echo=FALSE }

ggplot(drc, aes(x=year_id, y=value, color=source_type, shape=value_code)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Government Health Expenditure on Malaria, DRC") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))
ggplot(drc[value_code=='fs_malaria_domestic_public'], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (fs_malaria_domestic_public)") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))
#ggplot(drc[value_code=='func_malaria_cure_care'], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (func_malaria_cure_care)")
#ggplot(drc[value_code=='func_malaria_fprev'], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (func_malaria_prev)")

ggplot(drc[value_code=='fs_malaria_domestic_public'], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (fs_malaria_domestic_public)") + geom_smooth(method="lm") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))

ggplot(drc_ghe, aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, DRC (fs_malaria_domestic_public)") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))


```

The unique values in the data are `r unique(drc[, value_code])`. 
The unique sources in the data are `r unique(drc[, source_type])`

And the years we have each one for are 
```{r, echo=FALSE}
unique(drc[, .(year_id, value_code)])[order(year_id)]
unique(drc[, .(year_id, source_type)])[order(year_id)]
```


Some more specific analyses on GHE for malaria specifically `(value_code == fs_domestic_malaria_public)`
```{r, echo=FALSE}

ggplot(drc_ghe[source_type == "nha"], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, NHAs only") + geom_smooth(method="lm") + 
  scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))
ggplot(drc_ghe[source_type == "gfatm p" | source_type == "gfatm c"], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, Global Fund sources") + geom_smooth(method="lm") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))

ggplot(drc_ghe[source_type == "WMR"], aes(x=year_id, y=value, color=source_type, shape=source_type)) + geom_point() + theme_bw() + labs(x="Year", y="USD", title = "Domestic Public Spending on Malaria, other data sources") + geom_smooth(method="lm") + scale_y_continuous(limits=c(0, 30000000), labels=seq(0, 30000000, by=10000000))



```