---
title: "Senegal Performance Indicators"
author: "Francisco Rios Casas"
date: "10/29/2019"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(data.table)
library(ggplot2)
library(knitr)
library(dplyr)
library(kableExtra)
library(scales)

# set parameters of data to be analyzed
country = "sen"
main_indicators = c('impact_outcome_indicators_main', 'coverage_indicators_main')
recent_pudrs = c('SEN-Z-MOH_Progress  Report_30Jun2019   02 09 2019.xlsx', 
              'SEN-M-PNLP_Progress Report_S1 2019 Version finale du 15 Aout 2019.xlsx', 
              'PU-SEN-H-CNLS-S1-2019_15082019_finale.xlsx', 
              'SEN H ANCS PU (Jan-Juin19), LFA 5Sept19.xlsm')
LFA_verified_grants = c("SEN-Z-MOH PUDR (Juil-Dec18) LFA 19Avr19 MAJ 25apr19.xlsx",
                        "SEN M PNLP PUDR (Juil-Dec18) LFA, 10Mai19.xlsx",
                        "SEN H ANCS PU (Jan-Juin19), LFA 5Sept19.xlsm",
                        "SEN-H-CNLS PUDR (Juil-Dec18) LFA, 9Avr19.xlsx") # file names of most-recent LFA-verified grants

# load data
data <- readRDS("J:/Project/Evaluation/GF/outcome_measurement/multi_country/performance_indicators/pudr_indicator_extraction/cleaned_data/kpi_data_for_analyses.RDS")

# subset as appropriate
DT <- data
DT = DT[loc_name==country]
DT = DT[pudr_sheet %in% main_indicators]
DT = DT[file_name %in% recent_pudrs]
```


## 1. Introduction  

This document highlights the data completeness, data sources, and changes over time for Outcome Indicators in Senegal. Data for indicators all came from the latest available Progress Update and Disbursement Request forms (PUDRs) indicated below:  


```{r toc}
kable(recent_pudrs, col.names = c("Most Recent PUDRs available")) %>%
  kable_styling(bootstrap_options = "striped", position = "center")
```

## 2. Data Completeness  

For each Outcome Indicator, the grant Principal Recipient (PR) *usually* must establish an indicator "Baseline Value". This value--when available--can serve as a benchmark for progress reported in subsequent time periods. **However, in the Senegal grants, the use of the baseline value seems to be inconsistently applied as typically around 50% and 75% of indicators in a given PUDR do not include a Baseline Value.**

```{r completeness}
# Emily's code on {PUDR} Completeness

# Data
# What percentage of targets do they report on?
# Convert NAs to "Unknown"


# Convert NAs to "Unknown"
DT$plot_baseline <-NA
DT$plot_baseline[is.na(DT$baseline_value)] <- "Baseline Not Reported"
DT$plot_baseline[!is.na(DT$baseline_value)] <- "Baseline Value Reported"

ggplot(DT, aes(x=grant, fill=plot_baseline))+
  geom_bar(position = "fill")+
  labs(title="Proportion of PUDRs that report on baseline values", y="")+
  theme_bw(base_size = 10)+
  theme(legend.title=element_blank())+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))
```

In order to calculate the achievement of an Outcome Indicator for a given grant, it is necessary to have information on (1) the target that was originally set and (2) the result. Only the malaria (SEN-M-PNLP) and tuberculosis grant (SEN-Z-MOH) in Senegal provide sufficient data to understand how well the grants are meeting the goals that were originally established.  

```{r completeness2}
# create completeness indicator for baseline seperately from target and result

ggplot(DT, aes(x = grant, fill = completeness_rating)) +
  geom_bar(position = "fill")+
  labs(title="Completeness of Indicators according to grants", y="")+
  theme_bw(base_size = 10)+
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))+
  theme(legend.title=element_blank())
```

## 3. Data Sources  

"Data Sources" are listed in the PUDR to indicate where information reported to the global fund came from. Examples of data sources include reports from the disease-specific National Programs, surveys such as the Demographic Health Survey (DHS), and Health Management Information systems such as DHIS2.  

In some cases, the source of information used to establish the target of a goal is different from the source used to provide updates on progress. **In Senegal, the HIV grant to the National HIV Program (CNLS) had indicator sources change the most often: 12 of the 18 indicators in the grant changed sources.**  


```{r sources}
# create indicator variable of when result and target source differ
# subset data for checking changing indicators
table3 <-DT[,list(indicators=length(indicator_code), total_different=sum(sources_different, na.rm = TRUE)), by=c('grant')]

kable(table3, col.names = c("Grant", "Indicators", "Different Sources between Baseline and Result")) %>%
  column_spec(2, width = "10em") %>%
  column_spec(3, width = "10em") %>%
  kable_styling(bootstrap_options = "striped", position = "center")

```

  
**In the CNLS grant, the Baseline Values were established using reports from the National HIV Program, while the Result Values were obtained from a combination of DHIS2 data, indicators reported directly from the clinics providing services, and from other reports created by the Ministry of Health.**   


```{r sources_appendix1, include=FALSE}
# subset data for checking changing indicators
table3b <-DT[grant=="SEN-H-CNLS",.(indicator_code, brief_description_code, baseline_source, pr_result_source)]

kable(table3b, col.names = c("Indicator", "Description", "Baseline Source", "Result source"))


```

```{r sources_appendix2, include=FALSE}
table3c <-DT[grant=="SEN-Z-MOH",.(indicator_code, brief_description_code, baseline_source_code, pr_result_source_code, sources_different)]

kable(table3c, col.names = c("Indicator", "Description", "Baseline Source", "Result source", "indicx"))

```
  

## 4. Grant Results based on most recent PUDRs  


The "Achievement Ratio" is calculated in the PUDR to understand how close to the original target the Outcome Indicator result is. A value of 1.0 or greater means that the Indicators goal has been met or exceeded.


### 4.1. TB
```{r plot_results}
# plots grant performance in a loop for all grants in country
grants = unique(DT$grant)
plots = list()
i=1
for(g in grants) {
plots[[i]] = ggplot(DT[grant==g], aes(x=brief_description_code, y=ihme_result_achievement_ratio)) +
  geom_point() +
  labs(title=paste0(g, "\n performance"), y="Achievement Ratio", x="Indicator", caption=paste0("Source: ", unique(DT[grant==g]$file_name))) +
  geom_hline(yintercept = 1) +
  theme_bw()+
  coord_flip()+
  ylim(0,1.5)+
  theme(legend.position = "bottom")+
  theme_bw()
i=i+1
}
```

```{r}
plots[[4]]
```


**In the most recent PUDR, the TB grant has reported the lowest performance in the number of reporting units that had no stock-outs of anti-TB drugs. Number of MDR-TB patients that began second line treatment and health district coverage with TB services are the next indicators with the worst performance. Community referrals has also been an area where the indicator results reported have not been meeting the targets set. Section 6 below includes information on progress over time for some of these Indicators.** 


### 4.2. Malaria

```{r}
plots[[1]]
```


**The Malaria grant has been meeting all of its targets except with regards to the proportion of pregnant women who received at least three doses of intermittent preventive treatment (IPTp).**



### 4.3. HIV

```{r}
plots[[2]]
```

**The HIV ANCS grant has been performing lower than its targets in indicators that relate to sex workers, specifically. There are also a few Indicators with no updates available.**

```{r}
plots[[3]]
```
**The CNLS grant appears to have more indicators that are lower performing than the ANCS (NGO) grant.**

**The majority of Indicators between the two HIV grants are different, but in some cases--such as KP-3d(M) testing for HIV among people who inject drugs--both the CNLS and ANCS report different results for the same indicator. The PUDRs for each of the grants also list different targets: while the ANCS grant reports a higher value it also lists a higher target and the CNLS grant lists a lower result but also a much lower target leading to a much higher achievement ratio. Currently, only the ANCS PUDR is LFA-verified however.**  



## 5. Agreement between PR, LFA

After each PUDR is submitted by the PR, the LFA will verify the results reported on the Indicators. In some cases they may revise the original values provided by the PR.

The most recent LFA-verified PUDRs for this analysis include the following:


```{r}
kable(LFA_verified_grants, col.names = c("LFA Verified PUDR Files")) %>%
  kable_styling(bootstrap_options = "striped", position = "center")
```



```{r lfa_agreement}

data.agreement <- data[file_name%in%LFA_verified_grants & pr_result_value<=2000]

ggplot(data.agreement, aes(x=pr_result_value, y=lfa_result_value, color=grant))+
  geom_point()+
  labs(x="PR Reported Value", y="LFA Verified Values", title = paste0("Agreement between PR and LFA in grants"))+
  geom_abline(slope = 1, intercept = 0)+
  theme_bw()+
  scale_colour_discrete(name="Grants")
  #+ geom_text(data=data.agreement[pr_result_value>60 & lfa_result_value<50], aes(label=indicator_code), hjust="right", check_overlap = TRUE)

```


**In general, for the majority of indicators in the various grants, there was agreement between the values reported by the PR and those verified by the Global Fund LFA, except in the TB grant. In that case the LFA revised several of the reported results**


## 6. Results Over Time  

The tables below highlights how indicators available in the three most recent PUDRs have changed over time for each of the grants.
```{r}
# subset data

# subset parameters
variables_to_plot = c("ihme_result_achievement_ratio")

# subset to most recent pudrs and main_indicators
data2 <- data[loc_name==country & pudr_sheet%in%main_indicators]

# re-shape the data
data2 <- melt(data2, id.vars = c("indicator_code", "grant", "brief_description_code", "end_date_programmatic", "file_name", "type", "pudr_sheet"))
data2 <- data2[variable %in% variables_to_plot]

# change format
data2$value <- as.numeric(data2$value)
data2 <- data2[!is.na(data2$value),]

# subset data to only those indicators with three time points per grant
keep <- data2[,.N,by=indicator_code] #keep indicators with 3 time points
indicators <- keep[N==3,indicator_code] # create vector of indicators with 3 time point stop
data2 <- data2[indicator_code %in% indicators]

# plot the data
# write a loop to make plots for each grant
grants = unique(data2$grant)
dplots = list()
i=1
for(g in grants) {
dplots[[i]] = ggplot(data2[grant==g], aes(x=end_date_programmatic, y=value, group=brief_description_code, color=brief_description_code))+
  geom_point()+
  geom_line()+
  theme_bw()+
  scale_x_date(date_breaks = "6 month", 
                 labels=date_format("%b-%Y"))+
   labs(x="Date of PUDR", y="Achievement Ratio", title=paste0(g,"\n", "Performance Over Time"))+
  scale_colour_discrete(name="")+
  geom_hline(yintercept=1, linetype="dashed", color = "gray")+
  ylim(0, 1.5)+
  theme(legend.position="bottom", legend.text=element_text(size=6))+
  guides(col=guide_legend(ncol=1, keywidth = ))

i=i+1
}
```

```{r}
dplots[[1]]
```
**There has been an overall trend of improvement in indicators in the malaria grant, with the exception of IPTp provided to pregnant women which has not improved much since the first semester of 2018.**


```{r}
dplots[[2]]
```
**Indicators related to prevention programs and HIV testing among sex workers have also worsened since the first semester of 2018. While HIV testing among MSM has improved from where it was during the first semester of 2019, the target goal was not being met in the most recent PUDR.**


```{r}
dplots[[3]]
```
**The HIV-CNLS grant in Senegal has several indicator targets which have not been met and there has been little progress on the majority of them, with the exception of pregnant women who know their HIV status which despite lower performance throughout 2018 has been improving in the most recent period.**



```{r}
ggplot(data2[grant=="SEN-Z-MOH"], aes(x=end_date_programmatic, y=value, group=brief_description_code, color=brief_description_code))+
  geom_point()+
  geom_line()+
  theme_bw()+
  scale_x_date(date_breaks = "6 month", 
                 labels=date_format("%b-%Y"))+
   labs(x="Date of PUDR", y="Achievement Ratio", title=paste0("SEN-Z-MOH","\n", "Performance Over Time"))+
  scale_colour_discrete(name="")+
  geom_hline(yintercept=1, linetype="dashed", color = "gray")+
  ylim(0, 1.5)+
  theme(legend.position="right", legend.text=element_text(size=6))+
  guides(col=guide_legend(ncol=1, keywidth = ))
```

**MDR-TB identification and treatment (top two lines) is an area of large investment by the Global Fund in Senegal. Progress made in the year 2018 on both of those indicators has somewhat stalled in the first semester of 2019.** 

**TB cases identified through community referrals is also an important area of investment in Senegal; this indicator has remained just under-target in the most recent PUDRs despite making progress from being at 0 in early 2018 (bottom left line).** 
 

```{r, include=FALSE}
# target-setting over time

# subset parameters
trg_variables_to_plot = c("target_value")

# subset to most recent pudrs and main_indicators
data3 <- data[loc_name==country & pudr_sheet%in%main_indicators]

# re-shape the data
data3 <- melt(data3, id.vars = c("indicator_code", "grant", "brief_description_code", "end_date_programmatic", "file_name", "type", "pudr_sheet"))
data3 <- data3[variable %in% trg_variables_to_plot]

# change format
data3$value <- as.numeric(data3$value)
data3 <- data3[!is.na(data3$value),]

# subset data to only those indicators with three time points per grant
keeps <- data3[,.N,by=indicator_code] #keep indicators with 3 time points
trg_indicators <- keeps[N==3,indicator_code] # create vector of indicators with 3 time point stop
data3 <- data3[indicator_code %in% trg_indicators]

grants = unique(data3$grant)

cplots = list()
i=1
for(g in grants) {
cplots[[i]] = ggplot(data3[grant==g], aes(x=end_date_programmatic, y=value, group=brief_description_code, col=brief_description_code))+
  geom_point()+
  geom_line()+
  theme_bw()+
  facet_wrap(~variable)+
  scale_x_date(date_breaks = "6 month", 
                 labels=date_format("%b-%Y"))+
   labs(x="Date of PUDR", y="Target Value", title=paste0(g, "\n", "Target-setting Over Time"), caption = paste0("Data source: ", unique(data3$file_name)))+
  scale_colour_discrete(name="")+
  theme(legend.position="bottom")+
  guides(col=guide_legend(ncol=1))
  
i=i+1
}
cplots[[1]]
cplots[[2]]
cplots[[3]]
cplots[[4]]

```
