---
title: "Senegal Current Grant Overview"
author: "Emily Linebarger"
date: "May 10, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(data.table)
library(knitr)
library(ggplot2)
library(formattable)

budgets = readRDS("J:/Project/Evaluation/GF/resource_tracking/_gf_files_gos/sen/prepped_data/final_budgets.rds")
setDT(budgets)
expenditures = readRDS("J:/Project/Evaluation/GF/resource_tracking/_gf_files_gos/sen/prepped_data/final_expenditures.rds")
setDT(expenditures)

all_interventions = readRDS("C:/Users/elineb/Desktop/all_interventions.rds") #Pull in to get abbrev module. 
all_interventions = all_interventions[, .(module_eng, abbrev_mod_eng)] 
names(all_interventions) = c('gf_module', 'abbrev_mod')
all_interventions = unique(all_interventions)
all_interventions = all_interventions[!is.na(abbrev_mod)]

budgets = merge(budgets, all_interventions, by='gf_module', all.x=T)
expenditures = merge(expenditures, all_interventions, by='gf_module', all.x=T)
```
# Country overview 
There are currently four active grants in Senegal.   
1. SEN-H-ANCS (PR - Alliance Nationale des Communautés pour la Santé)  
2. SEN-H-CNLS (PR - Conseil National de Lutte contre le SIDA de la République du Sénégal)  
3. SEN-M-PNLP (PR - Ministry of Health and Social Action of the Republic of Senegal)   
4. SEN-Z-MOH (PR - Ministry of Health and Social Action of the Republic of Senegal), which is a combined TB/RSSH grant.   

# Funding landscape overview 
*All numbers reported in USD*
```{r Overview, echo=FALSE, warning = FALSE}
grant_summary = budgets[, .(budget=sum(budget, na.rm=T)), by=c('grant', 'grant_period')]
#Format budget as currency 
grant_summary[, budget:=currency(budget, digits = 0L)]
kable(grant_summary, caption="Total from detailed budgets")

absorption = expenditures[, .(budget=sum(budget, na.rm=T), expenditure=sum(expenditure, na.rm=T)), by='grant']
absorption[, absorption:=expenditure/budget]
absorption[, absorption:=round(absorption, 2)]
kable(absorption[, .(grant, budget, expenditure, absorption)], caption = "Reported expenditure and absorption")

```

```{r Analysis by module, echo=FALSE, warning=FALSE}
by_module = budgets[, .(budget=sum(budget, na.rm=T)), by=c('grant', 'abbrev_mod')]
grants = unique(budgets$grant)
for (var in grants){
  modplot = ggplot(by_module[grant==var], aes(x=reorder(abbrev_mod, budget), y=budget, fill=grant)) +
    geom_bar(stat="identity") +
    theme_bw() +
    theme(legend.title=element_blank()) +
    coord_flip() +
    labs(x="Module", y="Budget (USD)", title=paste0("Modules ranked by budget"))
  print(modplot)
}
```
For the modules with the highest spend in each grant, what are the main interventions? 
```{r Key Interventions, echo=FALSE, warning=FALSE}

sen_m_pnlp = budgets[grant=="SEN-M-PNLP" & abbrev_mod=="Vector control", .(budget=sum(budget, na.rm=T)), by=c('gf_intervention')]
kable(sen_m_pnlp[order(-budget)], caption="SEN-M-PNLP, Module: Vector Control")


sen_h_ancs = budgets[grant=="SEN-H-ANCS" & abbrev_mod=="Program mgmt", .(budget=sum(budget, na.rm=T)), by=c('gf_intervention')]
kable(sen_h_ancs[order(-budget)], caption="SEN-H-ANCS, Module: Program Management")

sen_h_cnls = budgets[grant=="SEN-H-CNLS" & abbrev_mod=="Treatment, care & support", .(budget=sum(budget, na.rm=T)), by=c('gf_intervention')]
kable(sen_h_cnls[order(-budget)], caption="SEN-H-CNLS, Module: Treatment, care & support")

sen_z_moh = budgets[grant=="SEN-Z-MOH" & abbrev_mod=="Care & prevention", .(budget=sum(budget, na.rm=T)), by=c('gf_intervention')]
kable(sen_h_cnls[order(-budget)], caption="SEN-Z-MOH, Module: Care & prevention")


```
