---
title: "PNLS Supply Chain Analysis-Double Check Test Kits"
output: beamer_presentation
---

```{r setup, include=FALSE, echo=FALSE}
rm(list=ls())
# -----------------------------------------------
# Set up R
#------------------------------------------------
library(data.table)
library(raster)
library(ggplot2)
library(ggrepel)
library(geosphere)
library(knitr)
library(tibble)
library(dplyr)
library(RColorBrewer)
library(rgdal)
library(rgeos)
library(rmarkdown)
library(maptools)
repo_root = "C:/Users/elineb/Documents/gf/" #Set to the root of your repository 
setwd(repo_root)
source('./core/standardizeDPSNames.R')

#Set up directories 
j = ifelse(Sys.info()[1]=='Windows', 'J:', '/home/j')
dir = paste0(j,  '/Project/Evaluation/GF/outcome_measurement/cod/dhis_data/prepped/pnls_final/') #Home directory
saveDir = paste0(j, 'Project/Evaluation/GF/outcome_measurement/cod/dhis_data/outputs/pnls/')
codeDir = paste0(repo_root, "outcome_measurement/all/cod/dhis/dhis_analysis/pnls_supplychain/")

source(paste0(codeDir, "4_set_up_data.r"))
source(paste0(codeDir, "4_pnls_supplychain_graph_functions.r"))
#-----------------------------------------------
#SET THE GLOBAL VARIABLES FOR THIS R MARKDOWN 
#-----------------------------------------------
drug_id = "k3JmmwNHkmY"
drug_name = "Double Check test kit"
full_date_range = "Jan 2017 - Dec 2018" #Imported into the graphs automatically - update as you get more data. 
partial_date_range = "Same time period: January - September" #Date range of 'AnnualDT' below - to have comparable time periods. Update as you get more data. 

```

## Reporting Completeness

```{r reporting1, echo=FALSE, warning=FALSE}
# Run for Determine test kits, but could also run for all drugs? 
plot = gen_report1(dt, drug_id, drug_name, full_date_range)
plot
```

## Reporting Completeness
```{r reporting2, echo=FALSE, warning=FALSE}
plot = gen_report2(dt, drug_id, drug_name, full_date_range)
plot
```

## Reporting completeness

```{r reporting3, echo=FALSE, warning=FALSE}
plot = gen_report3(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Stockouts of first-line test kits 
```{r stockout1, echo=FALSE, warning=FALSE}
plot = gen_stockout1(dt, drug_id, drug_name)
plot
```

## Percentage of facilities stocked out of first-line test kits

```{r stockout2, echo=FALSE, warning=FALSE}
plot = gen_stockout2(dt, drug_id, drug_name)
plot
```

## Total days stocked out by facility

```{r so_days1, echo=FALSE, warning=FALSE}
plot = gen_so_days1(dt, drug_id, drug_name, full_date_range)
plot
``` 

## Map of total facility-days of stockouts

```{r so_map1, echo=FALSE, warning=FALSE}
plot = gen_so_map1(dt, drug_id, drug_name, full_date_range)
plot
```

## Map of mean number of days stocked out

```{r so_map2, echo=FALSE, warning=FALSE}
plot = gen_so_map2(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Rate of change of stockouts

```{r roc1, echo=FALSE, warning=FALSE}
plot = gen_roc1(dt, drug_id, drug_name, full_date_range)
plot
```

## Stock-out rate of change, by month

```{r monthly_so_roc1, echo=FALSE, warning=FALSE}
plot = gen_monthly_so_roc1(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Districts with more stockouts in 2018 than 2017

```{r roc3, echo=FALSE, warning=FALSE}
plot = gen_roc3(dt, drug_id, drug_name, full_date_range)
plot
```

## Percentage of facility-days stocked out

```{r so_pct1, echo=FALSE, warning=FALSE}
plot = gen_so_pct1(dt, drug_id, drug_name, full_date_range)
plot
```

## Percentage of facility-days stocked out, with labels

```{r so_pct2, echo=FALSE, warning=FALSE}
plot = gen_so_pct2(dt, drug_id, drug_name, full_date_range)
plot
```

## Days stocked out by facility level (all)

```{r scatter1, echo=FALSE, warning=FALSE}
plot = gen_scatter1(dt, drug_id, drug_name, full_date_range)
plot
``` 

## Days stocked out by facility level (grouped)

```{r scatter2, echo=FALSE, warning=FALSE}
plot = gen_scatter2(dt, drug_id, drug_name, full_date_range)
plot
```

## Days stocked out by facility level (grouped, by year)

```{r scatter3, echo=FALSE, warning=FALSE}
plot = gen_scatter3(dt, drug_id, drug_name, full_date_range)
plot
```

## Stock-out duration, 2017 (No stockouts - 2 MO)

```{r cat1_2017, echo=FALSE, warning=FALSE}
plot = gen_cat1_2017(dt, coord_ann, drug_id, drug_name)
plot
``` 

## Stock-out duration, 2017 (2-month intervals)

```{r cat3_2017, echo=FALSE, warning=FALSE}
plot = gen_cat3_2017(dt, coord_ann, drug_id, drug_name)
plot
```

## Stock-out duration, 2018 (No stockouts - 2 MO)

```{r cat1_2018, echo=FALSE, warning=FALSE}
plot = gen_cat1_2018(dt, coord_ann, drug_id, drug_name)
plot
```

## Stock-out duration, 2018 (2-month intervals)

```{r cat3_2018, echo=FALSE, warning=FALSE}
plot = gen_cat3_2018(dt, coord_ann, drug_id, drug_name)
plot
```

## Drug units per facility, by district

```{r mean_units1, echo=FALSE, warning=FALSE}
plot = gen_mean_units1(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Drug units per facility by district, with labels 

```{r mean_units2, echo=FALSE, warning=FALSE}
plot = gen_mean_units2(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Facilities with a full month of stock-out, 2017

```{r full_months_17_1, echo=FALSE, warning=FALSE}
plot = gen_full_months1_17(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Proportion of facilities stocked out for a whole month, 2017

```{r full_months_17_2, echo=FALSE, warning=FALSE}
plot = gen_full_months2_17(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Facilities with a full month of stock-out, 2018

```{r full_months_18_1, echo=FALSE, warning=FALSE}
plot = gen_full_months1_18(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Proportion of facilities stocked out for a whole month, 2018

```{r full_months_18_2, echo=FALSE, warning=FALSE}
plot = gen_full_months2_18(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```
