---
title: "PNLS Supply Chain Analysis-First Line Treatment (TDF/3TC/EFV)"
output: beamer_presentation
---

```{r setup, include=FALSE, echo=FALSE}
# -----------------------------------------------
# Set up R
#------------------------------------------------
library(data.table)
library(raster)
library(ggplot2)
library(ggrepel)
library(geosphere)
library(knitr)
library(tibble)
library(dplyr)
library(RColorBrewer)
library(rgdal)
library(rgeos)
library(rmarkdown)
library(maptools)
repo_root = "C:/Users/elineb/Documents/gf/" #Set to the root of your repository 
setwd(repo_root)
source('./core/standardizeDPSNames.R')

#Set up directories 
j = ifelse(Sys.info()[1]=='Windows', 'J:', '/home/j')
dir = paste0(j,  '/Project/Evaluation/GF/outcome_measurement/cod/dhis_data/prepped/pnls_final/') #Home directory
saveDir = paste0(j, 'Project/Evaluation/GF/outcome_measurement/cod/dhis_data/outputs/pnls/')
codeDir = paste0(repo_root, "outcome_measurement/all/cod/dhis/dhis_analysis/pnls_supplychain/")

source(paste0(codeDir, "4_set_up_data.r"))
source(paste0(codeDir, "4_pnls_supplychain_graph_functions.r"))
#-----------------------------------------------
#SET THE GLOBAL VARIABLES FOR THIS R MARKDOWN 
#-----------------------------------------------
drug_id = "jJuipTLZK4o"
drug_name = "TDF/3TC/EFV"
full_date_range = "Jan 2017 - Dec 2018" #Imported into the graphs automatically - update as you get more data. 
partial_date_range = "Same time period: January - September" #Date range of 'AnnualDT' below - to have comparable time periods. Update as you get more data. 

```

## DRC Treatment Regimen
TDF+3TC+EFV is first-line treatment, and TDF+3TC+NVP, AZT+3TC+EFV, and AZT+3TC+NVP are second-line. We only have AZT+3TC+NVP available in the data, so this is the only drug in the graphs for second-line. 

## Reporting Completeness

```{r reporting1, echo=FALSE, warning=FALSE}
plot = gen_report1(dt, drug_id, drug_name, full_date_range)
plot
```

## Reporting Completeness
```{r reporting2, echo=FALSE, warning=FALSE}
plot = gen_report2(dt, drug_id, drug_name, full_date_range)
plot
```

## Reporting completeness

```{r reporting3, echo=FALSE, warning=FALSE}
plot = gen_report3(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Stockouts of first-line treatment drugs
```{r stockout1, echo=FALSE, warning=FALSE}
plot = gen_stockout1(dt, drug_id, drug_name)
plot
```

## Percentage of facilities stocked out of first-line treatment drugs

```{r stockout2, echo=FALSE, warning=FALSE}
plot = gen_stockout2(dt, drug_id, drug_name)
plot
```

## Total days stocked out by facility

```{r so_days1, echo=FALSE, warning=FALSE}
plot = gen_so_days1(dt, drug_id, drug_name, full_date_range)
plot
``` 


```{r so_map1, echo=FALSE, warning=FALSE, include=FALSE}
## Map of total facility-days of stockouts
plot = gen_so_map1(dt, drug_id, drug_name, full_date_range)
plot
```

## Map of mean number of days stocked out

```{r so_map2, echo=FALSE, warning=FALSE}
plot = gen_so_map2(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Rate of change of stockouts

```{r roc1, echo=FALSE, warning=FALSE}
plot = gen_roc1(dt, drug_id, drug_name, full_date_range)
plot
```


```{r monthly_so_roc1, echo=FALSE, warning=FALSE, include=FALSE}
## Stock-out rate of change, by month
plot = gen_monthly_so_roc1(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Districts with more stockouts in 2018 than 2017

```{r roc3, echo=FALSE, warning=FALSE}
plot = gen_roc3(dt, drug_id, drug_name, full_date_range)
plot
```


```{r so_pct1, echo=FALSE, warning=FALSE, include=FALSE}
## Percentage of facility-days stocked out
plot = gen_so_pct1(dt, drug_id, drug_name, full_date_range)
plot
```

## Percentage of facility-days stocked out, with labels

```{r so_pct2, echo=FALSE, warning=FALSE}
plot = gen_so_pct2(dt, drug_id, drug_name, full_date_range)
plot
```

## DPS stocked out for 85% of the year or greater
```{r so_pct3, echo=FALSE, warning=FALSE}
 stock = dt[element_id==drug_id & stock_category=="number_of_days_stocked_out", .(total_days=sum(expected_days, na.rm=T)), by=c('year', 'id', 'dps')]#Count all facilities that reported stock out information for this drug, with greater than 0 days in the month. 
  stock_add = dt[element_id==drug_id & stock_category=="number_of_days_stocked_out", .(days_out=sum(value, na.rm=T)), by=c('year', 'id', 'dps')]
  stock = merge(stock, stock_add, by=c('year', 'id', 'dps'), all=T)
  stock[, percent_out:=round(100*(days_out/total_days), 1)]
  
  #Pull DPS that have higher than 85% 
  high_so = unique(stock[percent_out>85, .(dps)])
  high_so_out = dt[dps%in%high_so$dps & !is.na(value) & stock_category=='number_of_days_stocked_out', .(facilities_reporting=length(unique(org_unit_id))), by=c('dps', 'year')]
  high_so_out2 = dt[dps%in%high_so$dps & !is.na(value) & stock_category=='number_of_days_stocked_out', .(org_unit_id, value, expected_days, level)]
  high_so_out2[, fac_pct_days_so:=(value/expected_days)*100]
  high_so_out2 = merge(high_so_out2, unique(dt[, .(org_unit_id, dps)]), by='org_unit_id')
  
  #Scatter this by level and DPS
  scatter3 = ggplot(high_so_out2[!is.na(level)], aes(x=level, y=fac_pct_days_so)) +
      geom_jitter(width=0.25, alpha=0.2) + theme_bw() +
    theme(axis.text.x = element_text(angle = 30, hjust=1)) + 
    facet_wrap(~dps)+
      labs(title=paste0('Percentage of facility days stocked out of ', drug_name, ' by facility,\nfor DPS with 85% of days stocked out or greater'), subtitle=full_date_range, x='Unique facilities',
           y='% Days stocked out')
  
  scatter3
 
```

## Number of facilities reporting for each DPS with 85% of days stocked out, by year

```{r, scatter4, echo=FALSE, warning=FALSE}
kable(high_so_out[order(dps, year)])
```

## Days stocked out by facility level (all)

```{r scatter1, echo=FALSE, warning=FALSE}
plot = gen_scatter1(dt, drug_id, drug_name, full_date_range)
plot
``` 


```{r scatter2, echo=FALSE, warning=FALSE, include=FALSE}
## Days stocked out by facility level (grouped)
plot = gen_scatter2(dt, drug_id, drug_name, full_date_range)
plot
```


```{r scatter3, echo=FALSE, warning=FALSE, include=FALSE}
## Days stocked out by facility level (grouped, by year)
plot = gen_scatter3(dt, drug_id, drug_name, full_date_range)
plot
```

## Prolonged stock-outs, 2017 (No stockouts - 2 MO)

```{r cat1_2017, echo=FALSE, warning=FALSE}
plot = gen_cat1_2017(dt, coord_ann, drug_id, drug_name)
plot
``` 

## Prolonged stock-outs, 2017 (2-month intervals)

```{r cat3_2017, echo=FALSE, warning=FALSE}
plot = gen_cat3_2017(dt, coord_ann, drug_id, drug_name)
plot
```

## Prolonged stock-outs, 2018 (No stockouts - 2 MO)

```{r cat1_2018, echo=FALSE, warning=FALSE}
plot = gen_cat1_2018(dt, coord_ann, drug_id, drug_name)
plot
```

## Prolonged stock-outs, 2018 (2-month intervals)

```{r cat3_2018, echo=FALSE, warning=FALSE}
plot = gen_cat3_2018(dt, coord_ann, drug_id, drug_name)
plot
```

## Drug units per facility, by district

```{r mean_units1, echo=FALSE, warning=FALSE}
plot = gen_mean_units1(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Drug units per facility by district, with labels

```{r mean_units2, echo=FALSE, warning=FALSE}
plot = gen_mean_units2(dt, coord_ann, drug_id, drug_name, full_date_range)
plot
```

## Facilities with a full month of stock-out, 2017

```{r full_months_17_1, echo=FALSE, warning=FALSE}
plot = gen_full_months1_17(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Proportion of facilities stocked out for a whole month, 2017

```{r full_months_17_2, echo=FALSE, warning=FALSE}
plot = gen_full_months2_17(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Facilities with a full month of stock-out, 2018

```{r full_months_18_1, echo=FALSE, warning=FALSE}
plot = gen_full_months1_18(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## Proportion of facilities stocked out for a whole month, 2018

```{r full_months_18_2, echo=FALSE, warning=FALSE}
plot = gen_full_months2_18(dt, coord_months, drug_id, drug_name, full_date_range)
plot
```

## EXTRA SLIDES 

## Reporting completeness, for stock-out data only 
```{r, extra1, echo=FALSE, warning=FALSE}
FacsReportingByDPS = dt[stock_category=='number_of_days_stocked_out', .(facilities_by_dps=length(unique(org_unit_id))), by=c('year', 'dps')]
  report_by_dps = merge(FacsReportingByDPS, dt[element_id==drug_id & stock_category=="number_of_days_stocked_out"], by=c('year', 'dps'))
  report_by_dps = unique(report_by_dps[, .(year, id, facilities_by_dps, dps)]) 
  report_map = merge(report_by_dps, coord_ann, by = c('id', 'year'), all.y = TRUE, allow.cartesian = TRUE)
  
  #----------------------
  #For each DPS, find it's center. 
  districts = unique(report_map$dps)
  districts = districts[!is.na(districts)]
  all_centers = data.table()
  for (district in districts){
    centers = report_map[dps==district, .(long, lat)]
    center = as.data.table(centroid(centers))
    center[, dps:=district]
    all_centers = rbind(all_centers, center)
  }
  
  # Generate a labels dataset
  labels = unique(report_map[, .(id, dps, facilities_by_dps, year)])
  labels[, label:= paste0(dps, ": ", facilities_by_dps)]
  labels = merge(labels, all_centers, by=c('dps'))
  #----------------------
  
  # mean_units2 = ggplot(kits_per_facility, aes(x=long, y=lat, group=group, fill=kits_per_fac)) + 
  #   geom_polygon() + 
  #   theme_void() + 
  #   facet_wrap(~year) + 
  #   scale_fill_gradientn('Kits per\nFacility', colours=ratio_colors) + 
  #   coord_fixed(ratio=1) + 
  #   scale_x_continuous('', breaks = NULL) + 
  #   scale_y_continuous('', breaks = NULL) + 
  #   labs(title=paste0(drug_name, " kits per facility, by district"), subtitle="Annual data restricted to Jan-Aug",
  #        caption="*Denominator only includes facilities with 'available, usable stock' of this drug",
  #        fill="Kits per facility") + 
  #   theme(plot.title=element_text(vjust=-1), plot.caption=element_text(vjust=6)) + 
  #   geom_label_repel(data = labels, aes(label = label, x = lon, y = lat, group = label), inherit.aes=FALSE, size=3)
  
  report_map1 = ggplot(report_map, aes(x=long, y=lat, group=group, fill=facilities_by_dps)) + 
    coord_fixed() +
    geom_polygon() + 
    geom_path(size=0.01) + 
    scale_fill_gradientn(colors=(brewer.pal(9, 'Blues'))) + 
    theme_void() +
    facet_wrap(~year, strip.position = "bottom") +
    labs(title=paste0("Reporting completeness* by year and district for ", drug_name), caption="Facilities reporting stock-out information", 
         fill="# of facilities reporting")+
    geom_label_repel(data = labels, aes(label = label, x = lon, y = lat, group = label), inherit.aes=FALSE, size=3)
  report_map1
```
