---
title: "Activity Classification Review"
author: "Emily Linebarger"
date: "July 23, 2019"
output: pdf_document
classoption: landscape
fontsize: 10pt
---

```{r setup, echo=FALSE, warning=FALSE}
# --------------------------------------------
# AUTHOR: Emily Linebarger
# PURPOSE: Review a few key examples of activity classifications 
# That aren't mapped correctly to GF modular framework. 
# DATE: July 22, 2019 
#----------------------------------------------

rm(list=ls())
library(data.table) 
library(knitr)
library(kableExtra)
library(openxlsx)

#Read in file 
dt = readRDS("J:/Project/Evaluation/GF/resource_tracking/_gf_files_gos/combined_prepped_data/budget_pudr_iterations.rds")
dt = dt[data_source=="fpm"]

#Do some fancy NLP string splitting - create a dictionary. 


# ---------------------------
# KVP 
#----------------------------
msm = dt[grep("msm|hsh", tolower(activity_description)), .(grant, grant_period, activity_description, orig_module, orig_intervention, gf_module, gf_intervention)]
msm = unique(msm)
#Search for errors by excluding the MSM module 
msm = msm[gf_module!="Comprehensive prevention programs for men who have sex with men"]
msm[, count:=1]

# ---------------------------
# GeneXpert
#----------------------------

xpert = dt[grep("xpert", tolower(activity_description)), .(grant, grant_period, orig_module, orig_intervention, gf_module, gf_intervention, activity_description, grant_disease)]
xpert = xpert[grant_disease%in%c('tb', 'rssh')]
xpert = unique(xpert)
xpert[, count:=1]

# ---------------------------
# Salaries 
#----------------------------
salary = dt[grep("salaire|salario|salary", tolower(activity_description))]
salary = unique(salary[, .(grant, grant_period, orig_module, orig_intervention, gf_module, gf_intervention, activity_description)])
salary = salary[gf_module!="Program management"]
#View(salaries[grant=="GTM-H-HIVOS", .(orig_module, orig_intervention, activity_description)])

#-----------------------------
# Program management - kind of a dead end so far. 
#-----------------------------
management = dt[grep("gestion|management|gestión", tolower(activity_description))]
management = unique(management[, .(grant, grant_period, orig_module, orig_intervention, gf_module, gf_intervention, activity_description)])
management = management[gf_module!="Program management"]

#--------------------------------
# Human Rights 
#--------------------------------
hr = dt[gf_module=="Programs to reduce human rights-related barriers to HIV services"]
hr = unique(hr[, .(grant, grant_period, orig_module, orig_intervention, gf_module, gf_intervention, activity_description)])

#Read in hand-processed sheet of these results 
hr_error = read.xlsx("I:/RTs_and_Projects/Evaluations/Global Fund PCE/Resource Tracking/NLP Examples 7.24.19.xlsx", sheet = "Example 3")
setDT(hr_error)
hr_error = hr_error[belongs==FALSE]

#--------------------------------
# Human Rights 
#--------------------------------
crs = dt[gf_module == "Community responses and systems"]
crs = unique(crs[, .(grant, grant_period, orig_module, orig_intervention, gf_module, gf_intervention, activity_description)])

#Can you go at this another way? Search for 'rights' and see what you turn up? EMILY 

#-----------------------------------------
# What are dead giveaway keywords?
#-----------------------------------------
# Surveys - DONE 
surveys = dt[grep("survey|enquete|encuesta", tolower(activity_description))]
surveys = unique(surveys[, .(grant, grant_period, orig_module, orig_intervention, gf_module, gf_intervention, activity_description)])
surveys = surveys[gf_intervention!="Surveys"]

#-----------------------------------------
# Can you easily classify drug-buying? 
#-----------------------------------------
art = dt[grep("art|antiretroviral|anti-retroviral", tolower(activity_description))]
art = unique(art[, .(grant, grant_period, orig_module, orig_intervention, gf_module, gf_intervention, activity_description)])



```


## Finding 1: Classifications for similar activity descriptions are inconsistent among and between PRs. 
Activity descriptions that mentioned "GeneXpert" were classified in multiple different ways among PRs during different grant periods. 
By grant, grant period, and module: 

```{r ex1, echo=F}
kable(xpert[, .(classifications=sum(count)), by=c('grant', 'grant_period', 'gf_module')], booktab=T)
```

By module alone: 

```{r ex1.1, echo=F}
kable(xpert[, .(classifications=sum(count)), by=c('gf_module')], booktab=T)
```


There was a mix of classifications between MDR-TB, general TB Care and Prevention, and TB/HIV. However, when reviewing the activity descriptions, it's unclear why these distinctions were made. 

```{r ex1.2, echo=F}
kable(xpert[, .(gf_module, activity_description)][order(gf_module)], booktabs=T, longtable=T) %>%
  column_spec(1, width="10em")%>%
  column_spec(2, width="50em") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size=8) %>%
  row_spec(c(5, 6, 7, 41, 47:48, 58:60), bold = T, background = "yellow")
```

## Finding 2: Human rights is questionably applied as a module, especially through mixing human rights and key populations. 

The following table is a hand-reviewed selection of activity descriptions that were classified under the module "Programs to reduce human rights-related barriers to HIV services". These are cases I feel are questionably classified under this module. The highlighted items are cases where the activity might be better classified under a key populations module. 

```{r ex2, echo=FALSE}
kable(hr_error[, .(orig_module, orig_intervention, activity_description)], booktabs=T, longtable=T)%>%
  column_spec(1:2, width="10em") %>%
  column_spec(3, width="40em") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size=8)%>%
  row_spec(c(6:8, 9, 12, 13:15, 16), bold = T, background = "yellow") %>%
  row_spec(c(10, 11), bold = T, background = "orange")

```

## Finding 3: RSSH classifications are inconsistent between grants 

One example is how surveys of various kinds are classified. This table represents all activity descriptions with the word "survey" (in English, French, or Spanish) that were not classified under the RSSH module "Health management information system and monitoring and evaluation" and the intervention "Surveys".
```{r ex3, echo=FALSE}
kable(surveys[, .(orig_module, orig_intervention, activity_description)], booktabs=T, longtable=T)%>%
  column_spec(1:2, width="10em") %>%
  column_spec(3, width="40em") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size=8)%>%
  row_spec(c(9, 11, 12, 20, 21), bold = T, background = "yellow")
```

## Finding 4: Program management and salary costs may be incorrectly labeled as disease-specific modules, skewing perceptions of funding. 

The table below represents all cases where "salary" was mentioned (in English, French, and Spanish) in an activity descrition, but the module was NOT marked as program management. 

```{r ex4, echo=FALSE}
kable(salary[, .(orig_module, orig_intervention, activity_description)], booktabs=T, longtable=T)%>%
  column_spec(1:2, width="10em") %>%
  column_spec(3, width="40em") %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), font_size=8)
```

# FINDING STATEMENTS IN PROGRESS

## Finding 5: Module and intervention categories are not well structured to capture the true intention of activity funding (“breadth” of an intervention is too large) ***


## Finding 6: Subpopulations are not neatly split in activity descriptions *** 