---
title: "Resource Tracking Data Gaps"
author: "Emily Linebarger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, echo=FALSE, warning=FALSE}
library(data.table) 
library(ggplot2)
library(knitr)
library(kableExtra) 

current_year = year(Sys.Date())
current_quarter = quarter(Sys.Date())
current_date = current_year + ((current_quarter/4)-0.25)

#Read in current budget and expenditure files (For absorption we are not trying to make a complete time series, so this check is less relevant, 
#   but absorption will have all of the same files as expenditure)
budgets = readRDS(paste0(box, "tableau_data/final_budgets.rds"))
budgets[, date:=year+((quarter/4)-0.25)]
budgets = unique(budgets[, .(grant, loc_name, date, file_name)])
pudrs = readRDS(paste0(box, "tableau_data/final_expenditures.rds"))
pudrs[, quarter:=quarter(start_date)]
pudrs[, year:=year(start_date)]
pudrs[, date:=year+((quarter/4)-0.25)]
pudrs[, concat:=paste0(grant, "_", grant_period)]
pudrs = unique(pudrs[, .(grant, loc_name, date, grant_period, concat)])


#Read in Global Fund metadata, and format it to match pudr/budget data above. 
ip_periods = fread(paste0(dir, "_gf_files_gos/metadata/grant_agreement_implementation_periods_dataset_201963.csv"))
ip_periods = ip_periods[GeographicAreaName%in%c('Congo (Democratic Republic)', 'Guatemala', 'Uganda', 'Senegal'), .(GeographicAreaName, ComponentName, GrantAgreementNumber, IsActive, ImplementationPeriodStartDate, ImplementationPeriodEndDate)]
names(ip_periods) <- c('loc_name', 'disease', 'grant', 'is_active', 'start_date', 'end_date') #Only need implementation period variables here! 
date_vars = c('start_date', 'end_date')
for (var in date_vars){
  ip_periods[nchar(get(var))==20, (var):=substr(get(var), 1, 8)]
  ip_periods[nchar(get(var))==21, (var):=substr(get(var), 1, 9)]
  ip_periods[nchar(get(var))==22, (var):=substr(get(var), 1, 10)]
  
  ip_periods[, (var):=as.Date(get(var), format="%m/%d/%Y")]
}

ip_periods[loc_name=="Congo (Democratic Republic)", loc_name:='cod']
ip_periods[loc_name=="Guatemala", loc_name:='gtm']
ip_periods[loc_name=="Senegal", loc_name:='sen']
ip_periods[loc_name=="Uganda", loc_name:='uga']

current_grants = ip_periods[is_active==TRUE]

# Need a dataset of expected grants and grant periods. 
expected_periods = ip_periods[, .(loc_name, grant, start_date, end_date)]

#Add variables. 
expected_periods[, start_year:=year(start_date)]
expected_periods[, end_year:=year(end_date)]
expected_periods[, start_quarter:=(quarter(start_date)/4)-0.25] 
expected_periods[, end_quarter:=(quarter(end_date)/4)-0.25] 
expected_periods[, start_decimal:=start_year+start_quarter]
expected_periods[, end_decimal:=end_year+end_quarter]

all_expected = data.table()
for (i in 1:nrow(expected_periods)){
  seq = seq(expected_periods[i, start_decimal], expected_periods[i, end_decimal], by=0.25)
  seq_table = data.table(date=seq)
  append_cols = expected_periods[i, .(grant, loc_name, start_year, end_year, start_quarter, end_quarter)]
    for (col in names(append_cols)){
      seq_table[, (col):=append_cols[, get(col)]]
    }  
  all_expected = rbind(all_expected, seq_table, fill=TRUE)
}

#Only worry about files from 2015 onwards; we have GMS data before then. 
all_expected = all_expected[date>=2015]

#Append Senegal files that we know we should have, but don't yet. 
pudrs = merge(pudrs, all_expected, all.y=T, by=c('loc_name', 'grant', 'date'))
pudrs[is.na(concat) & date<current_date, missing_pudr_date:=date]
missing_pudr_qtrs = unique(pudrs[is.na(concat) & date<=current_date, .(loc_name, grant, date)])
missing_pudr_qtrs[, quarter:=((date%%1)*4)+1]
missing_pudr_qtrs[, year:=floor(date)]

budgets = merge(budgets, all_expected, all.y=T, by=c('loc_name', 'grant', 'date'))
budgets[is.na(file_name) & date<current_date, missing_budget_date:=date]
missing_budget_qtrs = unique(budgets[is.na(file_name), .(loc_name, grant, date)])
missing_budget_qtrs[, quarter:=((date%%1)*4)+1]
missing_budget_qtrs[, year:=floor(date)]

#Save this output data
write.csv(pudrs, paste0(box, "tableau_data/missing_pudr_coverage.csv"), row.names=F)
write.csv(budgets, paste0(box, "tableau_data/missing_budget_coverage.csv"), row.names=F)

write.csv(pudrs, paste0(box, "tableau_data/archive/_missing_pudr_coverage_", Sys.Date(), ".csv"), row.names=F)
write.csv(budgets, paste0(box, "tableau_data/archive/missing_budget_coverage_", Sys.Date(), ".csv"), row.names=F)

pudr_coverage = function(country, country_name){
  pudr_coverage = ggplot(pudrs[loc_name==country]) + 
  geom_point(aes(x=date, y=grant), shape=19) + 
  geom_point(aes(x=missing_pudr_date, y=grant), color="blue", size=6, alpha=0.5, shape=15) +
  theme_bw() + 
  labs(title=paste0("PUDR coverage for ", country_name), subtitle="Black circles show expected grant quarters,\nand colored squares show missing data", x="Date", y="Grant", caption="*Only for grants active during 2015-2020")
  return(pudr_coverage)
}

budget_coverage = function(country, country_name){
  budget_coverage = ggplot(budgets[loc_name==country]) + 
  geom_point(aes(x=date, y=grant), shape=19) + 
  geom_point(aes(x=missing_budget_date, y=grant), color="red", size=6, alpha=0.5, shape=15) +
  theme_bw() + 
  labs(title=paste0("Budget coverage for ", country_name), subtitle="Black circles show expected grant quarters,\nand colored squares show missing data", x="Date", y="Grant", caption="*Only for grants active during 2015-2020")
  return(budget_coverage)
}


```

This report verifies that we have budget and PUDR files for all grants in the years between 2015-2020.  

The list of grants in this time period is:  
```{r, echo=FALSE, warning=FALSE}
kable(current_grants) %>%
  kable_styling(font_size=9)
```

This list is built from the metadata from the Global Fund's website.  

## Total missing quarters
The total number of missing budget quarters we have with this version of the database is: `r nrow(budgets[!is.na(missing_budget_date)])`
The total number of missing PUDR quarters we have with this version of the database is: `r nrow(pudrs[!is.na(missing_pudr_date)])`

## DRC Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
drc_pudrs = pudr_coverage("cod", "DRC")
drc_pudrs
kable(missing_pudr_qtrs[loc_name=="cod", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```

## DRC Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
drc_budget = budget_coverage("cod", "DRC")
drc_budget
kable(missing_budget_qtrs[loc_name=="cod", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```

## Guatemala Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
gtm_pudrs = pudr_coverage("gtm", "Guatemala")
gtm_pudrs
kable(missing_pudr_qtrs[loc_name=="gtm", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```

## Guatemala Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
gtm_budget = budget_coverage("gtm", "Guatemala")
gtm_budget
kable(missing_budget_qtrs[loc_name=="gtm", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```

## Senegal Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
sen_pudrs = pudr_coverage("sen", "Senegal")
sen_pudrs
kable(missing_pudr_qtrs[loc_name=="sen", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```

## Senegal Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
sen_budget = budget_coverage("sen", "Senegal")
sen_budget
kable(missing_budget_qtrs[loc_name=="sen", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```

## Uganda Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
uga_pudrs = pudr_coverage("uga", "Uganda")
uga_pudrs
kable(missing_pudr_qtrs[loc_name=="uga", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```

## Uganda Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
uga_budget = budget_coverage("uga", "Uganda")
uga_budget
kable(missing_budget_qtrs[loc_name=="uga", .(grant, year, quarter)]) %>%
  kable_styling(font_size=9)
```


