---
title: "Resource Tracking Data Gaps"
author: "Emily Linebarger"
date: "June 3, 2019"
output: pdf_document
---

```{r setup, echo=FALSE, warning=FALSE}
rm(list=ls())
library(data.table) 
library(ggplot2)
library(knitr)

# Current date 
current_date = Sys.Date()
current_year = year(current_date)
current_quarter = (quarter(current_date)/4)-0.25
current_date = current_year+current_quarter

#Read in data - only need to check budgets and expenditure files for now, alhough it might be nice to add in expenditure (?) EKL 6/12/19
dir = "J:/Project/Evaluation/GF/resource_tracking/_gf_files_gos/combined_prepped_data/"
budgets = readRDS(paste0(dir, "final_budgets.rds"))
budgets[, quarter:=(quarter/4)-0.25]
budgets[, date:=year+quarter]
budgets = unique(budgets[, .(grant, file_name, date, loc_name)])

#Because final expenditure data is NOT disaggregated into quarters; remerge this data with GOS from the budget_pudr_iterations file in the same wa as 2d_gf_aggregate_files.r
pudrs = readRDS(paste0(dir, "budget_pudr_iterations.rds"))
pudrs = pudrs[data_source=="pudr" & file_iteration=="final"]
gos_data = readRDS("J:/Project/Evaluation/GF/resource_tracking/_gf_files_gos/gos/prepped_data/prepped_gos_data.rds")
  
pudr_subset = pudrs[year>=2018]
gos_data = gos_data[year<2018]
pudr_subset = rbind(gos_data, pudr_subset, use.names=T, fill=T)

#Do the same transformations as the budget dataset
pudr_subset[, quarter:=quarter(start_date)]
pudr_subset[, year:=year(start_date)]
pudr_subset[, quarter:=(quarter/4)-0.25]
pudr_subset[, date:=year+quarter]
pudr_subset = unique(pudr_subset[ , .(grant, file_name, date, loc_name)])

#Read in Global Fund metadata, and format it to match pudr/budget data above. 
ip_periods = fread("J:/Project/Evaluation/GF/resource_tracking/_gf_files_gos/metadata/grant_agreement_implementation_periods_dataset_201963.csv")
ip_periods = ip_periods[GeographicAreaName%in%c('Congo (Democratic Republic)', 'Guatemala', 'Uganda', 'Senegal'), .(GeographicAreaName, ComponentName, GrantAgreementNumber, 
                                                                                                                      IsActive, 
                                                                                                                 ImplementationPeriodStartDate, ImplementationPeriodEndDate)]
names(ip_periods) <- c('loc_name', 'disease', 'grant', 'is_active', 'start_date', 'end_date') #Only need implementation period variables here! 
date_vars = c('start_date', 'end_date')
for (var in date_vars){
  ip_periods[nchar(get(var))==20, (var):=substr(get(var), 1, 8)]
  ip_periods[nchar(get(var))==21, (var):=substr(get(var), 1, 9)]
  ip_periods[nchar(get(var))==22, (var):=substr(get(var), 1, 10)]
  
  ip_periods[, (var):=as.Date(get(var), format="%m/%d/%Y")]
}

ip_periods[loc_name=="Congo (Democratic Republic)", loc_name:='cod']
ip_periods[loc_name=="Guatemala", loc_name:='gtm']
ip_periods[loc_name=="Senegal", loc_name:='sen']
ip_periods[loc_name=="Uganda", loc_name:='uga']

current_grants = ip_periods[is_active==TRUE]

# Need a dataset of expected grants and grant periods. 
expected_periods = ip_periods[, .(loc_name, grant, start_date, end_date)]

#Add variables. 
expected_periods[, start_year:=year(start_date)]
expected_periods[, end_year:=year(end_date)]
expected_periods[, start_quarter:=(quarter(start_date)/4)-0.25] 
expected_periods[, end_quarter:=(quarter(end_date)/4)-0.25] 
expected_periods[, start_decimal:=start_year+start_quarter]
expected_periods[, end_decimal:=end_year+end_quarter]

all_expected = data.table()
for (i in 1:nrow(expected_periods)){
  seq = seq(expected_periods[i, start_decimal], expected_periods[i, end_decimal], by=0.25)
  seq_table = data.table(date=seq)
  append_cols = expected_periods[i, .(grant, loc_name, start_year, end_year, start_quarter, end_quarter)]
    for (col in names(append_cols)){
      seq_table[, (col):=append_cols[, get(col)]]
    }  
  all_expected = rbind(all_expected, seq_table, fill=TRUE)
}

#Only worry about files from 2015 onwards; we have GMS data before then. 
all_expected = all_expected[date>=2015]

#Append Senegal files that we know we should have, but don't yet. 
pudr_subset = merge(pudr_subset, all_expected, all.y=T, by=c('loc_name', 'grant', 'date'))
pudr_subset[is.na(file_name) & date<current_date, missing_pudr_date:=date]
missing_pudr_qtrs = unique(pudr_subset[is.na(file_name) & date<=current_date, .(loc_name, grant, date)])
missing_pudr_qtrs[, quarter:=((date%%1)*4)+1]
missing_pudr_qtrs[, year:=floor(date)]

budgets = merge(budgets, all_expected, all.y=T, by=c('loc_name', 'grant', 'date'))
budgets[is.na(file_name) & date<current_date, missing_budget_date:=date]
missing_budget_qtrs = unique(budgets[is.na(file_name), .(loc_name, grant, date)])
missing_budget_qtrs[, quarter:=((date%%1)*4)+1]
missing_budget_qtrs[, year:=floor(date)]

pudr_coverage = function(country, country_name){
  pudr_coverage = ggplot(pudr_subset[loc_name==country]) + 
  geom_point(aes(x=date, y=grant), shape=19) + 
  geom_point(aes(x=missing_pudr_date, y=grant), color="blue", size=6, alpha=0.5, shape=15) +
  theme_bw() + 
  labs(title=paste0("PUDR coverage for ", country_name), subtitle="Black circles show expected grant quarters,\nand colored squares show missing data", x="Date", y="Grant", caption="*Only for grants active during 2015-2020")
  return(pudr_coverage)
}

budget_coverage = function(country, country_name){
  budget_coverage = ggplot(budgets[loc_name==country]) + 
  geom_point(aes(x=date, y=grant), shape=19) + 
  geom_point(aes(x=missing_budget_date, y=grant), color="red", size=6, alpha=0.5, shape=15) +
  theme_bw() + 
  labs(title=paste0("Budget coverage for ", country_name), subtitle="Black circles show expected grant quarters,\nand colored squares show missing data", x="Date", y="Grant", caption="*Only for grants active during 2015-2020")
  return(budget_coverage)
}


```

This report verifies that we have budget and PUDR files for all grants in the years between 2015-2020.  

The list of grants in this time period is:  
```{r, echo=FALSE, warning=FALSE}
kable(current_grants)
```

This list is built from the metadata from the Global Fund's website.  

## DRC Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
drc_pudrs = pudr_coverage("cod", "DRC")
drc_pudrs
kable(missing_pudr_qtrs[loc_name=="cod", .(grant, year, quarter)])
```

## DRC Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
drc_budgets = budget_coverage("cod", "DRC")
drc_budgets
kable(missing_budget_qtrs[loc_name=="cod", .(grant, year, quarter)])
```

## Guatemala Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
gtm_pudrs = pudr_coverage("gtm", "Guatemala")
gtm_pudrs
kable(missing_pudr_qtrs[loc_name=="gtm", .(grant, year, quarter)])
```

## Guatemala Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
gtm_budgets = budget_coverage("gtm", "Guatemala")
gtm_budgets
kable(missing_budget_qtrs[loc_name=="gtm", .(grant, year, quarter)])
```

## Senegal Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
sen_pudrs = pudr_coverage("sen", "Senegal")
sen_pudrs
kable(missing_pudr_qtrs[loc_name=="sen", .(grant, year, quarter)])
```

## Senegal Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
sen_budgets = budget_coverage("sen", "Senegal")
sen_budgets
kable(missing_budget_qtrs[loc_name=="sen", .(grant, year, quarter)])
```

## Uganda Missing PUDRs
```{r, warning=FALSE, echo=FALSE}
uga_pudrs = pudr_coverage("uga", "Uganda")
uga_pudrs
kable(missing_pudr_qtrs[loc_name=="uga", .(grant, year, quarter)])
```

## Uganda Missing Budgets 
```{r, warning=FALSE, echo=FALSE}
uga_budgets = budget_coverage("uga", "Uganda")
uga_budgets
kable(missing_budget_qtrs[loc_name=="uga", .(grant, year, quarter)])
```


