---
title: "Senegal 2018-2020 budget revisions"
author: "Emily Linebarger"
date: "October 10, 2019"
output: pdf_document
classoption: landscape
---

```{r setup, echo=FALSE, warning=FALSE}
library(data.table)
library(ggplot2)
library(knitr)
library(readxl)
library(scales)
options(scipen=100)

dt = readRDS("C:/Users/elineb/Box Sync/Global Fund Files/SEN/prepped_data/budget_revisions.rds") #This is at the intervention, year, quarter level. 
setnames(dt, c('gf_module', 'gf_intervention'), c('module_eng', 'intervention_eng'))
all_mods = readRDS("J:/Project/Evaluation/GF/resource_tracking/modular_framework_mapping/all_interventions.rds")
all_mods = unique(all_mods[, .(module_eng, intervention_eng, abbrev_mod_eng, abbrev_int_eng)])
dt = merge(dt, all_mods, by=c('module_eng', 'intervention_eng'), all.x=T)

#Collapse to module level. 
by_module = dt[, .(v0=sum(v0, na.rm=T), v1=sum(v1, na.rm=T)), by=c('grant', 'grant_period', 'abbrev_mod_eng')]
by_module = melt(by_module, id.vars=c('grant', 'grant_period', 'abbrev_mod_eng'), value.name="budget", variable.name="version")
by_module[, label:=paste0("$", round(budget))]
by_module[, version:=factor(version, levels=c('v1', 'v0'), labels=c('v1', 'v0'))] #Reverse this factoring so it shows up on the graph in the right order. 

#Collapse to intervention level. 
by_intervention = dt[, .(v0=sum(v0, na.rm=T), v1=sum(v1, na.rm=T)), by=c('grant', 'grant_period', 'abbrev_mod_eng', 'abbrev_int_eng')]
by_intervention = melt(by_intervention, id.vars=c('grant', 'grant_period', 'abbrev_mod_eng', 'abbrev_int_eng'), value.name="budget", variable.name="version")
by_intervention[, label:=paste0("$", round(budget))]
by_intervention[, version:=factor(version, levels=c('v1', 'v0'), labels=c('v1', 'v0'))] #Reverse this factoring so it shows up on the graph in the right order. 

#Read in the master file list to pull the names of the file revisions. 
file_list = data.table(read_excel("C:/Users/elineb/Box Sync/Global Fund Files/master_file_list.xlsx"))

```

# List of budget revisions for Senegal
```{r list_files, echo=FALSE, warning=FALSE}
revision_concats = unique(dt[, paste0(grant, "_", grant_period)])
file_list = unique(file_list[paste0(grant, "_", grant_period)%in%revision_concats & data_source=="budget" & 
                      file_iteration%in%c('final', 'revision'), .(grant, grant_period, file_name, update_date)])
file_list[, update_date:=as.Date(update_date, format="%Y-%m-%d")]
file_list = file_list[order(grant, grant_period, update_date)]

# Add an 'order' variable
file_list[, revision_order:=seq(0, 5, by=1), by=c('grant', 'grant_period')]
file_list[, version:=paste0("version ", revision_order)]
file_list$revision_order <- NULL
kable(file_list[, 1:3])
kable(file_list[, 4:5])
```

## SEN-H-ANCS module-level breakdown of revisions
```{r SEN_H_ANCS1, echo=FALSE, warning=FALSE}
ggplot(by_module[grant=="SEN-H-ANCS"], aes(x=abbrev_mod_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("SEN-H-ANCS budget revisions, by module"), x="Module", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

Program management is clearly the largest module. To better view the differences, the graph on the next page shows the module breakdown without program management.  

## SEN-H-ANCS module-level breakdown of revisions, without program management

```{r SEN_H_ANCS2, echo=FALSE, warning=FALSE}
ggplot(by_module[grant=="SEN-H-ANCS" & abbrev_mod_eng!="Program mgmt"], aes(x=abbrev_mod_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("SEN-H-ANCS budget revisions, by module"), x="Module", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

On this graph, it's easier to see that there is new funding for "TB/HIV" and "Human rights barriers" in the first revision, and more money has been added to prevention programs for MSM, IJU, and CSW and their clients. 

To dig a little deeper, the graphs on the next pages break down the changes to the modules "TB/HIV" and "Human rights barriers" at the intervention-level. 

## Revisions under the "TB/HIV" module
```{r SEN_H_ANCS3, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="SEN-H-ANCS" & abbrev_mod_eng=="TB/HIV"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("TB/HIV budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

New funding has been allocated to "Engaging all care providers."

## Revisions under the "Reducing human rights barriers" module

```{r SEN_H_ANCS4, echo=FALSE, warning=FALSE}
by_intervention[abbrev_int_eng=="Sensitization of lawmakers and law enforcement agents", abbrev_int_eng:="Sensitization of lawmakers \nand law enforcement agents"]

ggplot(by_intervention[grant=="SEN-H-ANCS" & abbrev_mod_eng=="Human rights barriers"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("Human rights budget revisions"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

New funding has been allocated to several new human rights interventions, the largest of which is "Stigma and discrimination reduction."

## SEN-H-CNLS module-level breakdown of revisions
```{r SEN_H_CNLS1, echo=FALSE, warning=FALSE}
ggplot(by_module[grant=="SEN-H-CNLS"], aes(x=abbrev_mod_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  # geom_text(data=by_module, aes(x=abbrev_mod_eng, y=budget, label=label), vjust=0) + 
  theme_bw() + 
  coord_flip() + 
  theme(axis.text.x=element_text(angle=30, vjust=0.5)) + 
  labs(title=paste0("SEN-H-CNLS budget revisions, by module"), x="Module", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)

```

This slide shows the module-level revisions for the CNLS HIV grant. There is new funding under the modules "Procurement and supply chain management (PSM)" and "Human rights barriers", and more funding has been added to several modules targeting key populations, just like the ANCS grant. On the next pages, I break down the intervention-level changes to the modules "Prevention programs for injecting drug users (IJU)", "Prevention programs for MSM", "Human rights barriers", and "Procurement and supply chain management".

## Revisions under the "Prevention for IJU" module
```{r SEN_H_CNLS3, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="SEN-H-CNLS" & abbrev_mod_eng=="Prevention programs for IJU"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("Prevention for injecting drug users"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

A large amount of money has been allocated to "Other interventions for injectable drug users". 

## Revisions under the "Prevention for MSM" module
```{r SEN_H_CNLS4, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="SEN-H-CNLS" & abbrev_mod_eng=="Prevention programs for MSM"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("Prevention for men who have sex with men"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

Again, a large amount of money has been allocated to the "other interventions" category. New funding has been added to prevention of co-infections, interventions for young MSM, and HIV testing for MSM. 

## Revisions under the "Reducing human rights barriers" module
```{r SEN_H_CNLS5, echo=FALSE, warning=FALSE}
ggplot(by_intervention[grant=="SEN-H-CNLS" & abbrev_mod_eng=="Human rights barriers"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("Reducing human rights barriers"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

There is new funding for several interventions aimed at reducing human rights barriers, the largest of which is "Stigma and discrimination reduction". 

## Revisions under the "Procurement and supply chain management" module
```{r SEN_H_CNLS6, echo=FALSE, warning=FALSE}
by_intervention[abbrev_int_eng=="Other procurement and supply chain management intervention(s)", abbrev_int_eng:="Other PSM interventions"]
ggplot(by_intervention[grant=="SEN-H-CNLS" & abbrev_mod_eng=="PSM"], aes(x=abbrev_int_eng, y=budget, fill=version)) + 
  geom_bar(stat="identity", position="dodge") + 
  theme_bw() + 
  coord_flip() + 
  labs(title=paste0("Procurement and supply chain management"), x="Intervention", y="Budget", fill="Version") + 
  scale_y_continuous(labels = scales::dollar)
```

There has been some money allocated to the "other" category within this module. 

